---
title: "Snake tracking"
author: "Marc Vez-Garzón"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
editor: visual
---

# Introduction

This document includes all the parts dedicated to analyse the movement and behaviour of multiple individuals of *Hemorrhois hippocrepis* collected on the island of Ibiza during the months of April to June 2023 and 2024.

In this study, we try to find if there are significant differences in the use of space, boldness, and exploratory behaviour between regions of the island recently invaded with populations of *Podarcis pityusensis* (endemic lizard of the island of Ibiza, threatened by this invasive snake) and regions of the island invaded for a long time and without natural populations of lizards.

# Packages

```{r packages}
#| warning: false
library(viridis)
library(readxl)
library(writexl)
library(ggplot2)
library(tidyr)
library(plyr)
library(sp)
library(lubridate)
library(dplyr)
library(irr)
library(utils)
library(tibble)
library(EnvStats)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(Hmisc)
library(ggpubr)
library(vegan)
library(sm)
library(survminer)
library(ggfortify)
library(factoextra)
library(graphics)
library(olsrr)
library(corrplot)
library(car)
library(censReg)
library(patchwork)
```

# Load data

Load the required files needed to run this script. Files can be found in https://github.com/marcvez/Snake_tracking.git

```{r data loading}
#| warning: false
# Set working directory
setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking")

# Load file with snake info
Snake_data <- read_excel("Snake_captures.xlsx")

# Subset tracked snakes
Snake_data_tracked <- Snake_data[Snake_data$Tracking %in% c("Y"), ]

# Subset tracked + genetics/isotopes + x-ray sample 
Snake_data_tracked <- Snake_data_tracked[!is.na(Snake_data_tracked$SVL), ]

# Load snake 3D array
load("Snake_array.Rda")

# Load arena 3D array
load("Arena_array.Rda")

# Load summary file from all snake trials
summary_tracking <- read_excel("summary_tracking.xlsx")

# Load summary file from all snake trials + environmental variables
summary_tracking_complete <- read_excel("summary_tracking_complete.xlsx")

```

For some analyses, we might need a data frame with only the first trials of each snakes, without gravid females, or with only those individuals that got out of the initial refuge and explored the arena.

```{r data filtering}
#| warning: false
time_limit <- 3000
# No repeated indiviuals, only with new ones
summary_tracking_complete_no_rep <- summary_tracking_complete[summary_tracking_complete$Repeatability == "N",]

# Also, for some individuals, we are only going to use individuals that actually got out of the refuge (Body-out != 3000)
summary_tracking_complete_no_rep_outside <- summary_tracking_complete_no_rep[summary_tracking_complete_no_rep$Time_body_out_sec_raw != time_limit, ]

# Create one without gravid individuals that got out (visual exploration)
summary_tracking_complete_no_rep_outside_noeggs <- summary_tracking_complete_no_rep_outside[summary_tracking_complete_no_rep_outside$N_eggs == 0, ]

# Create one without gravid individuals with censored data (head-out and body-out)
summary_tracking_complete_no_rep_noeggs <- summary_tracking_complete_no_rep[summary_tracking_complete_no_rep$N_eggs == 0, ]
```

## Snake position extraction

Don't run if `Snake_array` and `Arena_array` have been already loaded, as these functions take a lot of time to run.

`Snake_array`

```{r snake array}
#| warning: false
#| eval: false
# Variable names for each column
variable_names <- c("t", "x", "y", "Central_area", "Intermediate_area", "Refuge_area_1", "Refuge_area_2", "Initial_refuge", "Distant_wall", "Lateral_1", "Lateral_2", "Arena_area", "Area", "dist_hole_0", "dist_hole_1", "dist_hole_2", "dist_hole_3", "dist_hole_4", "dist_min", "hole_dist_min", "dist_traveled", "behaviour")

# We create empty data frame that is going to store: x axis = time step, y = variables, z = individuals
Snake_array <-  array(NA, dim = c(3600, length(variable_names), nrow(Snake_data_tracked)))

# We set the column names for the variables across the array
dimnames(Snake_array)[[2]] <- variable_names

# We set the ID code for all tracked snakes across the array
dimnames(Snake_array)[[3]] <- Snake_data_tracked$Snake_ID


# For each tracked individual 
# THIS TAKES A LOT OF TIME. DO NOT RUN AND LOAD DIRECTLY FROM WD
for(i in 1:nrow(Snake_data_tracked)){
  
  # If it is the first trial (no repeatability)...
  if(Snake_data_tracked$Repeatability[i] == "N"){
    
    # ... set the working directory in the correct folder...
    setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking/Excels tracking")
    
    # ... and if it didn't get out of the original refuge, fill the correspondent data layer with 0. Otherwise...
    if(Snake_data_tracked$Time_head_out_sec_raw[i] == "NA"){
      
      # We copy an already complete file
      coordinates_complete <- as.data.frame(Snake_array[, , 1])
      
      # And fill with 0
      coordinates_complete$x <- 0
      coordinates_complete$y <- 0
      coordinates_complete$Central_area <- 0
      coordinates_complete$Intermediate_area <- 0
      coordinates_complete$Refuge_area_1 <- 0
      coordinates_complete$Refuge_area_2 <- 0
      coordinates_complete$Initial_refuge <- 0
      coordinates_complete$Distant_wall <- 0
      coordinates_complete$Lateral_1 <- 0
      coordinates_complete$Lateral_2 <- 0
      coordinates_complete$Arena_area <- 0 
      coordinates_complete$Area <- "Initial refuge"
      coordinates_complete$dist_hole_0 <- 0
      coordinates_complete$dist_hole_1 <- 0.86
      coordinates_complete$dist_hole_2 <- 1.3
      coordinates_complete$dist_hole_3 <- 1.3
      coordinates_complete$dist_hole_4 <- 0.86
      coordinates_complete$dist_min <- 0
      coordinates_complete$hole_dist_min <- "dist_hole_0"
      coordinates_complete$dist_traveled <- 0
      coordinates_complete$behaviour <- "hidden"
      
      # It might happen that the snake is being recorded for more than 1h, and we want to know their behaviour in 1h-tests. It's for this reason that we are going to cut all videos to 1h (3600 sec). If the video lasts less than 1h, for the moment we are not going to do anything.
      if(nrow(coordinates_complete) > 3600){
        
        coordinates_complete <- coordinates_complete[1:3600, ]
        
      } else {
        
      }
      
      Snake_array[1:nrow(coordinates_complete), , i] <- as.matrix(coordinates_complete[ , ])
      
      assign("Snake_array", Snake_array, envir=globalenv())
      
      paste("Snake_array", i)
      
      
      # ... extract data from its tracked Excel file.
    } else if(Snake_data_tracked$Time_head_out_sec_raw[i] != "NA"){
      
      # Select data based on snake's ID
      data_name <- paste0(Snake_data_tracked$Snake_ID[i], "_coordinates.xlsx")
      
      # Standardized name for data
      coordinates <- read_excel(data_name)
      
      # Select arena limits and refuges from specific individual
      name_arena <- paste0(Snake_data_tracked$Snake_ID[i], "_arena.xlsx")
      
      arena <- read_excel(name_arena)
      
      # How much time does the snake hide before first head-out
      sec_bf_head_out <- as.numeric(Snake_data_tracked$Time_head_out_sec_raw[i])
      
      # Let's find which seconds is the snake missing in the recording (hidden)
      missing_time <- min(coordinates$t):max(coordinates$t)
      
      # How many seconds does the snake stay inside the refuges after coming out the initial refuge
      missing_time_recorded <- length(missing_time[!missing_time %in% coordinates$t])
      
      # Which seconds is the snake missing
      missing_time_sec <- missing_time[!missing_time %in% coordinates$t]
      
      
      # See in which area is the snake at each time step
      
      # The snake, even if it is outside the arena, is still over a certain area, so we are going to widen the limits of the areas for detecting in which area is the snake even if it is outside the arena limits. It does not have an effect over the behaviour
      
      off_limit <- 0.2
      
      # Central area
      Central_area <- point.in.polygon(coordinates$x, 
                                       coordinates$y, 
                                       c(arena[1, 2]- 0.75, arena[2, 2] - 0.75, arena[3, 2] + 0.75, arena[4, 2] + 0.75), 
                                       c(arena[1, 3] + 0.5, arena[2, 3] - 0.5, arena[3, 3] - 0.5, arena[4, 3] + 0.5))
      
      coordinates$Central_area <- Central_area
      
      # Intermediate area
      Intermediate_area <- point.in.polygon(coordinates$x, 
                                            coordinates$y, 
                                            c(arena[1, 2]- 0.5, arena[2, 2] - 0.5, arena[3, 2] + 0.5, arena[4, 2] + 0.5), 
                                            c(arena[1, 3] + 0.25, arena[2, 3] - 0.25, arena[3, 3] - 0.25, arena[4, 3] + 0.25))
      
      coordinates$Intermediate_area <- Intermediate_area
      
      coordinates$Intermediate_area <- coordinates$Intermediate_area - coordinates$Central_area
      
      # Refuge area 1
      Refuge_area_1 <- point.in.polygon(coordinates$x, 
                                        coordinates$y, 
                                        c(arena[1, 2]- 0.25, arena[2, 2] - 0.25, arena[2, 2] - 0.5, arena[2, 2] - 0.5),
                                        c(arena[1, 3] + 0.25, arena[2, 3] - 0.25, arena[2, 3] - 0.25, arena[1, 3] + 0.25))
      
      coordinates$Refuge_area_1 <- Refuge_area_1
      
      # Refuge area 2
      Refuge_area_2 <- point.in.polygon(coordinates$x, 
                                        coordinates$y, 
                                        c(arena[4, 2] + 0.5, arena[3, 2] + 0.5, arena[3, 2] + 0.25, arena[4, 2] + 0.25),
                                        c(arena[4, 3] + 0.25, arena[3, 3] - 0.25, arena[3, 3] - 0.25, arena[4, 3] + 0.25))
      
      coordinates$Refuge_area_2 <- Refuge_area_2
      
      
      # Initial refuge area
      Initial_refuge <- point.in.polygon(coordinates$x, 
                                         coordinates$y, 
                                         c(arena[1, 2] - 0.25, arena[1, 2] -0.25, arena[4, 2] + 0.25, arena[4, 2] + 0.25),
                                         c(arena[1, 3] - off_limit, arena[1, 3] + 0.25, arena[4, 3] + 0.25, arena[4, 3] - off_limit))
      
      coordinates$Initial_refuge <- Initial_refuge
      
      
      # Distant wall area
      Distant_wall <- point.in.polygon(coordinates$x, 
                                       coordinates$y, 
                                       c(arena[2, 2] - 0.25, arena[2, 2] - 0.25, arena[3, 2] + 0.25, arena[3, 2] + 0.25),
                                       c(arena[2, 3] - 0.25, arena[2, 3] + off_limit, arena[3, 3] + off_limit, arena[3, 3] - 0.25))
      
      coordinates$Distant_wall <- Distant_wall
      
      
      # Lateral refuge area 1
      Lateral_1 <- point.in.polygon(coordinates$x, 
                                    coordinates$y, 
                                    c(arena[1, 2] + off_limit, arena[2, 2] + off_limit, arena[2, 2] - 0.25, arena[1, 2] - 0.25),
                                    c(arena[1, 3] - off_limit, arena[2, 3] + off_limit, arena[2, 3] + off_limit, arena[1, 3] - off_limit))
      
      coordinates$Lateral_1 <- Lateral_1
      
      # Lateral refuge area 2
      Lateral_2 <- point.in.polygon(coordinates$x, 
                                    coordinates$y, 
                                    c(arena[3, 2] - off_limit, arena[4, 2] - off_limit, arena[4, 2] + 0.25, arena[3, 2] + 0.25),
                                    c(arena[3, 3] + off_limit, arena[4, 3] - off_limit, arena[4, 3] - off_limit, arena[3, 3] + off_limit))
      
      coordinates$Lateral_2 <- Lateral_2
      
      
      # Arena area. We are going to expand the arena limits 2cm on each side, in order to take into account edge exploration as inside the arena, not ouside
      Arena_area <- point.in.polygon(coordinates$x, 
                                     coordinates$y, 
                                     c(arena[1, 2] + 0.02, arena[2, 2] + 0.02, arena[3, 2] - 0.02, arena[4, 2] - 0.02),
                                     c(arena[1, 3] - 0.02, arena[2, 3] + 0.02, arena[3, 3] + 0.02, arena[4, 3] - 0.02))
      
      coordinates$Arena_area <- Arena_area
      
      # Column that indicates the area the snake is at every time step
      for(j in 1:nrow(coordinates)){
        
        if(coordinates$Arena_area[j] == 1 | coordinates$Arena_area[j] == 0){
          
          if(coordinates$Lateral_2[j] == 1){
            
            coordinates$Area[j] <- "Lateral 2"
            
          } else if(coordinates$Lateral_1[j] == 1){
            
            coordinates$Area[j] <- "Lateral 1"
            
          } else if(coordinates$Distant_wall[j] == 1){
            
            coordinates$Area[j] <- "Distant wall"
            
          } else if(coordinates$Refuge_area_1[j] == 1){
            
            coordinates$Area[j] <- "Refuge area 1"
            
          } else if(coordinates$Refuge_area_2[j] == 1) {
            
            coordinates$Area[j] <- "Refuge area 2"
            
          } else if(coordinates$Initial_refuge[j] == 1) {
            
            coordinates$Area[j] <- "Initial refuge"
            
          } else if(coordinates$Intermediate_area[j] == 1) {
            
            coordinates$Area[j] <- "Intermediate area"
            
          } else if(coordinates$Central_area[j] == 1){
            
            coordinates$Area[j] <- "Central area"
            
          } else  {
            
            if (j == 1){
              
              # Initial frame
              coordinates$Area[j] <- "Initial refuge"
              
            } else {
              
              # If any of the conditions are met, repeat previous area
              coordinates$Area[j] <- coordinates$Area[j - 1]
              
            }
            
          }
          
        }
        
      }
      
      
      
      # Distance to each hole
      
      # Entrance hole (hole_0)
      hole_0 <- c(arena[5, 2], arena[5, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_0[k] <- sqrt((coordinates$x[k] - as.numeric(hole_0[1]))^2 + (coordinates$y[k] - as.numeric(hole_0[2]))^2)
        
      }
      
      # Hole 1
      hole_1 <- c(arena[6, 2], arena[6, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_1[k] <- sqrt((coordinates$x[k] - as.numeric(hole_1[1]))^2 + (coordinates$y[k] - as.numeric(hole_1[2]))^2)
        
      }
      
      # Hole 2
      hole_2 <- c(arena[7, 2], arena[7, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_2[k] <- sqrt((coordinates$x[k] - as.numeric(hole_2[1]))^2 + (coordinates$y[k] - as.numeric(hole_2[2]))^2)
        
      }
      
      # Hole 3
      hole_3 <- c(arena[8, 2], arena[8, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_3[k] <- sqrt((coordinates$x[k] - as.numeric(hole_3[1]))^2 + (coordinates$y[k] - as.numeric(hole_3[2]))^2)
        
      }
      
      # Hole 4
      hole_4 <- c(arena[9, 2], arena[9, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_4[k] <- sqrt((coordinates$x[k] - as.numeric(hole_4[1]))^2 + (coordinates$y[k] - as.numeric(hole_4[2]))^2)
        
      }
      
      
      # Which is the hole the snake is closest to at each moment
      for(k in 1:nrow(coordinates)){
        
        a <- coordinates$dist_hole_0[k]
        b <- coordinates$dist_hole_1[k]
        c <- coordinates$dist_hole_2[k]
        d <- coordinates$dist_hole_3[k]
        e <- coordinates$dist_hole_4[k]
        
        # Write the distance to the closest hole
        coordinates$dist_min[k] <- min(c(a, b, c, d, e))
        
        # And the name of that hole
        coordinates$hole_dist_min[k] <- colnames(coordinates[which((coordinates[k, ] == coordinates$dist_min[k]) == TRUE)])[1]
        
      }
      
      
      # Which distance has the snake traveled from 1 sec to the next one
      dist_traveled <- data.frame(matrix(ncol = 1, nrow = nrow(coordinates)))
      
      for(k in 2:nrow(coordinates)){
        
        dist_traveled[k, 1] <- sqrt((coordinates$x[k] - coordinates$x[k - 1])^2 + (coordinates$y[k] - coordinates$y[k - 1])^2)
        
      }
      
      dist_traveled[1, 1] <- 0
      
      coordinates$dist_traveled <- as.numeric(dist_traveled[, 1])
      
      
      # Which behaviour is the snake exhibiting?
      for(k in 1:nrow(coordinates)){
        
        if(coordinates$Arena_area[k] == 0){
          
          if(coordinates$dist_traveled[k] < 0.01){
            
            coordinates$behaviour[k] <- "hiding / head-out"
            
          } else {
            
            coordinates$behaviour[k] <- "exploring wall"
            
          }
          
        } else {
          
          if(coordinates$dist_traveled[k] < 0.01 && coordinates$dist_min[k] < 0.1){
            
            coordinates$behaviour[k] <- "head-out"
            
          } else if(coordinates$dist_traveled[k] < 0.01 && coordinates$dist_min[k] > 0.1){
            
            coordinates$behaviour[k] <- "immobile exposed"
            
          } else {
            
            coordinates$behaviour[k] <- "moving"
            
          }
          
        }
        
      }
      
      
      # We put the correct time steps to each frame
      coordinates$t <- coordinates$t + as.numeric(sec_bf_head_out) + 1
      
      # We create a data set that is going to keep all the information to what happens during the whole video
      
      # How long is the total video (nº sec recorded + nº sec snake is hidden bf exiting for the first time + nº seconds hidden during recording)
      total_frames <- as.numeric(sec_bf_head_out) + nrow(coordinates) + missing_time_recorded
      
      # Column names
      col_names_coordinates <- colnames(coordinates)
      
      # We create the data frame
      coordinates_complete <- data.frame(matrix(ncol = ncol(coordinates), nrow = total_frames))
      
      # And put the column names
      colnames(coordinates_complete) <- col_names_coordinates
      
      # Seconds that the whole recordind lasts
      coordinates_complete$t <- seq(1:total_frames)
      
      # Now, we assign each recorded parameter to its correspondent second on the 1h record
      for(k in 1:nrow(coordinates)){
        
        values_coordinates <- coordinates[k, ]
        
        coordinates_complete[as.numeric(values_coordinates[1]), ] <- values_coordinates
        
        
      }
      
      
      # Values for when the snake is still inside the initial refuge
      for(j in 1:nrow(coordinates_complete)){
        
        # If the snake hasn't exited the initial refuge
        if(j <= as.numeric(sec_bf_head_out)){
          
          coordinates_complete$x[j] <- 0
          coordinates_complete$y[j] <- 0
          coordinates_complete$Central_area[j] <- 0
          coordinates_complete$Intermediate_area[j] <- 0
          coordinates_complete$Refuge_area_1[j] <- 0
          coordinates_complete$Refuge_area_2[j] <- 0
          coordinates_complete$Initial_refuge[j] <- 0
          coordinates_complete$Distant_wall[j] <- 0
          coordinates_complete$Lateral_1[j] <- 0
          coordinates_complete$Lateral_2[j] <- 0
          coordinates_complete$Arena_area[j] <- 0 
          coordinates_complete$Area[j] <- "Initial refuge"
          coordinates_complete$dist_hole_0[j] <- 0
          coordinates_complete$dist_hole_1[j] <- coordinates_complete$dist_hole_1[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_hole_2[j] <- coordinates_complete$dist_hole_2[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_hole_3[j] <- coordinates_complete$dist_hole_3[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_hole_4[j] <- coordinates_complete$dist_hole_4[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_min[j] <- 0
          coordinates_complete$hole_dist_min[j] <- "dist_hole_0"
          coordinates_complete$dist_traveled[j] <- 0
          coordinates_complete$behaviour[j] <- "hidden"
          
          
        } else {
          
          # If the snake is missing (hidden), which parameters are going to exhibit those frames. We are going to copy the parameters from the last recorded frame, except specific parameters.
          if(is.na(coordinates_complete$x[j]) == TRUE) {
            
            coordinates_complete$x[j] <- coordinates_complete$x[j - 1]
            coordinates_complete$y[j] <- coordinates_complete$y[j - 1]
            coordinates_complete$Central_area[j] <- coordinates_complete$Central_area[j - 1]
            coordinates_complete$Intermediate_area[j] <- coordinates_complete$Intermediate_area[j - 1]
            coordinates_complete$Refuge_area_1[j] <- coordinates_complete$Refuge_area_1[j - 1]
            coordinates_complete$Refuge_area_2[j] <- coordinates_complete$Refuge_area_2[j - 1]
            coordinates_complete$Initial_refuge[j] <-  coordinates_complete$Initial_refuge[j - 1]
            coordinates_complete$Distant_wall[j] <- coordinates_complete$Distant_wall[j - 1]
            coordinates_complete$Lateral_1[j] <- coordinates_complete$Lateral_1[j - 1]
            coordinates_complete$Lateral_2[j] <- coordinates_complete$Lateral_2[j - 1]
            coordinates_complete$Arena_area[j] <- coordinates_complete$Arena_area[j - 1]
            coordinates_complete$Area[j] <- coordinates_complete$Area[j - 1]
            coordinates_complete$dist_hole_0[j] <- coordinates_complete$dist_hole_0[j - 1]
            coordinates_complete$dist_hole_1[j] <- coordinates_complete$dist_hole_1[j - 1]
            coordinates_complete$dist_hole_2[j] <- coordinates_complete$dist_hole_2[j - 1]
            coordinates_complete$dist_hole_3[j] <- coordinates_complete$dist_hole_3[j - 1]
            coordinates_complete$dist_hole_4[j] <- coordinates_complete$dist_hole_4[j - 1]
            coordinates_complete$dist_min[j] <- 0
            coordinates_complete$hole_dist_min[j] <- coordinates_complete$hole_dist_min[j - 1]
            coordinates_complete$dist_traveled[j] <- 0
            coordinates_complete$behaviour[j] <- "hidden"
            
          }
          
        }
        
      }
      
      # It might happen that the snake is being recorded for more than 1h, and we want to know their behaviour in 1h-tests. It's for this reason that we are going to cut all videos to 1h (3600 sec). If the video lasts less than 1h, for the moment we are not going to do anything.
      if(nrow(coordinates_complete) > 3600){
        
        coordinates_complete <- coordinates_complete[1:3600, ]
        
      } else {
        
      }
      
      Snake_array[1:nrow(coordinates_complete), , i] <- as.matrix(coordinates_complete[ , ])
      
      assign("Snake_array", Snake_array, envir=globalenv())
      
      paste("Snake_array", i)
      
    }
    
    # If it is a snake used for repeatability, ...
  } else if(Snake_data_tracked$Repeatability[i] == "Y"){
    
    # ... set the correct wd ...
    setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking/Excels tracking/Replicates")
    
    # ... and check if the snake did get out of the initial refuge or not. If it didn't, fill with 0.
    if(Snake_data_tracked$Time_head_out_sec_raw[i] == "NA"){
      
      
      # We copy an already complete file
      coordinates_complete <- as.data.frame(Snake_array[, , 1])
      
      # And fill with 0
      coordinates_complete$x <- 0
      coordinates_complete$y <- 0
      coordinates_complete$Central_area <- 0
      coordinates_complete$Intermediate_area <- 0
      coordinates_complete$Refuge_area_1 <- 0
      coordinates_complete$Refuge_area_2 <- 0
      coordinates_complete$Initial_refuge <- 0
      coordinates_complete$Distant_wall <- 0
      coordinates_complete$Lateral_1 <- 0
      coordinates_complete$Lateral_2 <- 0
      coordinates_complete$Arena_area <- 0 
      coordinates_complete$Area <- "Initial refuge"
      coordinates_complete$dist_hole_0 <- 0
      coordinates_complete$dist_hole_1 <- 0.86
      coordinates_complete$dist_hole_2 <- 1.3
      coordinates_complete$dist_hole_3 <- 1.3
      coordinates_complete$dist_hole_4 <- 0.86
      coordinates_complete$dist_min <- 0
      coordinates_complete$hole_dist_min <- "dist_hole_0"
      coordinates_complete$dist_traveled <- 0
      coordinates_complete$behaviour <- "hidden"
      
      # It might happen that the snake is being recorded for more than 1h, and we want to know their behaviour in 1h-tests. It's for this reason that we are going to cut all videos to 1h (3600 sec). If the video lasts less than 1h, for the moment we are not going to do anything.
      if(nrow(coordinates_complete) > 3600){
        
        coordinates_complete <- coordinates_complete[1:3600, ]
        
      } else {
        
      }
      
      Snake_array[1:nrow(coordinates_complete), , i] <- as.matrix(coordinates_complete[ , ])
      
      assign("Snake_array", Snake_array, envir=globalenv())
      
      paste("Snake_array", i)
      
      
      
      # If the snake did get out of the refuge, extract data from excel
    } else if(Snake_data_tracked$Time_head_out_sec_raw[i] != "NA"){
      
      # Select data based on snake's ID
      data_name <- paste0(Snake_data_tracked$Snake_ID[i], "_coordinates.xlsx")
      
      # Standardized name for data
      coordinates <- read_excel(data_name)
      
      # Select arena limits and refuges from specific individual
      name_arena <- paste0(Snake_data_tracked$Snake_ID[i], "_arena.xlsx")
      
      arena <- read_excel(name_arena)
      
      # How much time does the snake hide before first head-out
      sec_bf_head_out <- as.numeric(Snake_data_tracked$Time_head_out_sec_raw[i])
      
      # Let's find which seconds is the snake missing in the recording (hidden)
      missing_time <- min(coordinates$t):max(coordinates$t)
      
      # How many seconds does the snake stay inside the refuges after coming out the initial refuge
      missing_time_recorded <- length(missing_time[!missing_time %in% coordinates$t])
      
      # Which seconds is the snake missing
      missing_time_sec <- missing_time[!missing_time %in% coordinates$t]
      
      
      # See in which area is the snake at each time step
      
      # The snake, even if it is outside the arena, is still over a certain area, so we are going to widen the limits of the areas for detecting in which area is the snake even if it is outside the arena limits. It does not have an effect over the behaviour
      
      off_limit <- 0.2
      
      # Central area
      Central_area <- point.in.polygon(coordinates$x, 
                                       coordinates$y, 
                                       c(arena[1, 2]- 0.75, arena[2, 2] - 0.75, arena[3, 2] + 0.75, arena[4, 2] + 0.75), 
                                       c(arena[1, 3] + 0.5, arena[2, 3] - 0.5, arena[3, 3] - 0.5, arena[4, 3] + 0.5))
      
      coordinates$Central_area <- Central_area
      
      # Intermediate area
      Intermediate_area <- point.in.polygon(coordinates$x, 
                                            coordinates$y, 
                                            c(arena[1, 2]- 0.5, arena[2, 2] - 0.5, arena[3, 2] + 0.5, arena[4, 2] + 0.5), 
                                            c(arena[1, 3] + 0.25, arena[2, 3] - 0.25, arena[3, 3] - 0.25, arena[4, 3] + 0.25))
      
      coordinates$Intermediate_area <- Intermediate_area
      
      coordinates$Intermediate_area <- coordinates$Intermediate_area - coordinates$Central_area
      
      # Refuge area 1
      Refuge_area_1 <- point.in.polygon(coordinates$x, 
                                        coordinates$y, 
                                        c(arena[1, 2]- 0.25, arena[2, 2] - 0.25, arena[2, 2] - 0.5, arena[2, 2] - 0.5),
                                        c(arena[1, 3] + 0.25, arena[2, 3] - 0.25, arena[2, 3] - 0.25, arena[1, 3] + 0.25))
      
      coordinates$Refuge_area_1 <- Refuge_area_1
      
      # Refuge area 2
      Refuge_area_2 <- point.in.polygon(coordinates$x, 
                                        coordinates$y, 
                                        c(arena[4, 2] + 0.5, arena[3, 2] + 0.5, arena[3, 2] + 0.25, arena[4, 2] + 0.25),
                                        c(arena[4, 3] + 0.25, arena[3, 3] - 0.25, arena[3, 3] - 0.25, arena[4, 3] + 0.25))
      
      coordinates$Refuge_area_2 <- Refuge_area_2
      
      
      # Initial refuge area
      Initial_refuge <- point.in.polygon(coordinates$x, 
                                         coordinates$y, 
                                         c(arena[1, 2] - 0.25, arena[1, 2] -0.25, arena[4, 2] + 0.25, arena[4, 2] + 0.25),
                                         c(arena[1, 3] - off_limit, arena[1, 3] + 0.25, arena[4, 3] + 0.25, arena[4, 3] - off_limit))
      
      coordinates$Initial_refuge <- Initial_refuge
      
      
      # Distant wall area
      Distant_wall <- point.in.polygon(coordinates$x, 
                                       coordinates$y, 
                                       c(arena[2, 2] - 0.25, arena[2, 2] - 0.25, arena[3, 2] + 0.25, arena[3, 2] + 0.25),
                                       c(arena[2, 3] - 0.25, arena[2, 3] + off_limit, arena[3, 3] + off_limit, arena[3, 3] - 0.25))
      
      coordinates$Distant_wall <- Distant_wall
      
      
      # Lateral refuge area 1
      Lateral_1 <- point.in.polygon(coordinates$x, 
                                    coordinates$y, 
                                    c(arena[1, 2] + off_limit, arena[2, 2] + off_limit, arena[2, 2] - 0.25, arena[1, 2] - 0.25),
                                    c(arena[1, 3] - off_limit, arena[2, 3] + off_limit, arena[2, 3] + off_limit, arena[1, 3] - off_limit))
      
      coordinates$Lateral_1 <- Lateral_1
      
      # Lateral refuge area 2
      Lateral_2 <- point.in.polygon(coordinates$x, 
                                    coordinates$y, 
                                    c(arena[3, 2] - off_limit, arena[4, 2] - off_limit, arena[4, 2] + 0.25, arena[3, 2] + 0.25),
                                    c(arena[3, 3] + off_limit, arena[4, 3] - off_limit, arena[4, 3] - off_limit, arena[3, 3] + off_limit))
      
      coordinates$Lateral_2 <- Lateral_2
      
      
      # Arena area. We are going to expand the arena limits 2cm on each side, in order to take into account edge exploration as inside the arena, not ouside
      Arena_area <- point.in.polygon(coordinates$x, 
                                     coordinates$y, 
                                     c(arena[1, 2] + 0.02, arena[2, 2] + 0.02, arena[3, 2] - 0.02, arena[4, 2] - 0.02),
                                     c(arena[1, 3] - 0.02, arena[2, 3] + 0.02, arena[3, 3] + 0.02, arena[4, 3] - 0.02))
      
      coordinates$Arena_area <- Arena_area
      
      # Column that indicates the area the snake is at every time step
      for(j in 1:nrow(coordinates)){
        
        if(coordinates$Arena_area[j] == 1 | coordinates$Arena_area[j] == 0){
          
          if(coordinates$Lateral_2[j] == 1){
            
            coordinates$Area[j] <- "Lateral 2"
            
          } else if(coordinates$Lateral_1[j] == 1){
            
            coordinates$Area[j] <- "Lateral 1"
            
          } else if(coordinates$Distant_wall[j] == 1){
            
            coordinates$Area[j] <- "Distant wall"
            
          } else if(coordinates$Refuge_area_1[j] == 1){
            
            coordinates$Area[j] <- "Refuge area 1"
            
          } else if(coordinates$Refuge_area_2[j] == 1) {
            
            coordinates$Area[j] <- "Refuge area 2"
            
          } else if(coordinates$Initial_refuge[j] == 1) {
            
            coordinates$Area[j] <- "Initial refuge"
            
          } else if(coordinates$Intermediate_area[j] == 1) {
            
            coordinates$Area[j] <- "Intermediate area"
            
          } else if(coordinates$Central_area[j] == 1){
            
            coordinates$Area[j] <- "Central area"
            
          } else  {
            
            if (j == 1){
              
              # Initial frame
              coordinates$Area[j] <- "Initial refuge"
              
            } else {
              
              # If any of the conditions are met, repeat previous area
              coordinates$Area[j] <- coordinates$Area[j - 1]
              
            }
            
          }
          
        }
        
      }
      
      
      
      # Distance to each hole
      
      # Entrance hole (hole_0)
      hole_0 <- c(arena[5, 2], arena[5, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_0[k] <- sqrt((coordinates$x[k] - as.numeric(hole_0[1]))^2 + (coordinates$y[k] - as.numeric(hole_0[2]))^2)
        
      }
      
      # Hole 1
      hole_1 <- c(arena[6, 2], arena[6, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_1[k] <- sqrt((coordinates$x[k] - as.numeric(hole_1[1]))^2 + (coordinates$y[k] - as.numeric(hole_1[2]))^2)
        
      }
      
      # Hole 2
      hole_2 <- c(arena[7, 2], arena[7, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_2[k] <- sqrt((coordinates$x[k] - as.numeric(hole_2[1]))^2 + (coordinates$y[k] - as.numeric(hole_2[2]))^2)
        
      }
      
      # Hole 3
      hole_3 <- c(arena[8, 2], arena[8, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_3[k] <- sqrt((coordinates$x[k] - as.numeric(hole_3[1]))^2 + (coordinates$y[k] - as.numeric(hole_3[2]))^2)
        
      }
      
      # Hole 4
      hole_4 <- c(arena[9, 2], arena[9, 3])
      
      for(k in 1:nrow(coordinates)){
        
        coordinates$dist_hole_4[k] <- sqrt((coordinates$x[k] - as.numeric(hole_4[1]))^2 + (coordinates$y[k] - as.numeric(hole_4[2]))^2)
        
      }
      
      
      # Which is the hole the snake is closest to at each moment
      for(k in 1:nrow(coordinates)){
        
        a <- coordinates$dist_hole_0[k]
        b <- coordinates$dist_hole_1[k]
        c <- coordinates$dist_hole_2[k]
        d <- coordinates$dist_hole_3[k]
        e <- coordinates$dist_hole_4[k]
        
        # Write the distance to the closest hole
        coordinates$dist_min[k] <- min(c(a, b, c, d, e))
        
        # And the name of that hole
        coordinates$hole_dist_min[k] <- colnames(coordinates[which((coordinates[k, ] == coordinates$dist_min[k]) == TRUE)])[1]
        
      }
      
      
      # Which distance has the snake traveled from 1 sec to the next one
      dist_traveled <- data.frame(matrix(ncol = 1, nrow = nrow(coordinates)))
      
      for(k in 2:nrow(coordinates)){
        
        dist_traveled[k, 1] <- sqrt((coordinates$x[k] - coordinates$x[k - 1])^2 + (coordinates$y[k] - coordinates$y[k - 1])^2)
        
      }
      
      dist_traveled[1, 1] <- 0
      
      coordinates$dist_traveled <- as.numeric(dist_traveled[, 1])
      
      
      # Which behaviour is the snake exhibiting?
      for(k in 1:nrow(coordinates)){
        
        if(coordinates$Arena_area[k] == 0){
          
          if(coordinates$dist_traveled[k] < 0.01){
            
            coordinates$behaviour[k] <- "hiding / head-out"
            
          } else {
            
            coordinates$behaviour[k] <- "exploring wall"
            
          }
          
        } else {
          
          if(coordinates$dist_traveled[k] < 0.01 && coordinates$dist_min[k] < 0.1){
            
            coordinates$behaviour[k] <- "head-out"
            
          } else if(coordinates$dist_traveled[k] < 0.01 && coordinates$dist_min[k] > 0.1){
            
            coordinates$behaviour[k] <- "immobile exposed"
            
          } else {
            
            coordinates$behaviour[k] <- "moving"
            
          }
          
        }
        
      }
      
      
      # We put the correct time steps to each frame
      coordinates$t <- coordinates$t + as.numeric(sec_bf_head_out) + 1
      
      # We create a data set that is going to keep all the information to what happens during the whole video
      
      # How long is the total video (nº sec recorded + nº sec snake is hidden bf exiting for the first time + nº seconds hidden during recording)
      total_frames <- as.numeric(sec_bf_head_out) + nrow(coordinates) + missing_time_recorded
      
      # Column names
      col_names_coordinates <- colnames(coordinates)
      
      # We create the data frame
      coordinates_complete <- data.frame(matrix(ncol = ncol(coordinates), nrow = total_frames))
      
      # And put the column names
      colnames(coordinates_complete) <- col_names_coordinates
      
      # Seconds that the whole recordind lasts
      coordinates_complete$t <- seq(1:total_frames)
      
      # Now, we assign each recorded parameter to its correspondent second on the 1h record
      for(k in 1:nrow(coordinates)){
        
        values_coordinates <- coordinates[k, ]
        
        coordinates_complete[as.numeric(values_coordinates[1]), ] <- values_coordinates
        
        
      }
      
      
      # Values for when the snake is still inside the initial refuge
      for(j in 1:nrow(coordinates_complete)){
        
        # If the snake hasn't exited the initial refuge
        if(j <= as.numeric(sec_bf_head_out)){
          
          coordinates_complete$x[j] <- 0
          coordinates_complete$y[j] <- 0
          coordinates_complete$Central_area[j] <- 0
          coordinates_complete$Intermediate_area[j] <- 0
          coordinates_complete$Refuge_area_1[j] <- 0
          coordinates_complete$Refuge_area_2[j] <- 0
          coordinates_complete$Initial_refuge[j] <- 0
          coordinates_complete$Distant_wall[j] <- 0
          coordinates_complete$Lateral_1[j] <- 0
          coordinates_complete$Lateral_2[j] <- 0
          coordinates_complete$Arena_area[j] <- 0 
          coordinates_complete$Area[j] <- "Initial refuge"
          coordinates_complete$dist_hole_0[j] <- 0
          coordinates_complete$dist_hole_1[j] <- coordinates_complete$dist_hole_1[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_hole_2[j] <- coordinates_complete$dist_hole_2[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_hole_3[j] <- coordinates_complete$dist_hole_3[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_hole_4[j] <- coordinates_complete$dist_hole_4[as.numeric(sec_bf_head_out) + 1]
          coordinates_complete$dist_min[j] <- 0
          coordinates_complete$hole_dist_min[j] <- "dist_hole_0"
          coordinates_complete$dist_traveled[j] <- 0
          coordinates_complete$behaviour[j] <- "hidden"
          
          
        } else {
          
          # If the snake is missing (hidden), which parameters are going to exhibit those frames. We are going to copy the parameters from the last recorded frame, except specific parameters.
          if(is.na(coordinates_complete$x[j]) == TRUE) {
            
            coordinates_complete$x[j] <- coordinates_complete$x[j - 1]
            coordinates_complete$y[j] <- coordinates_complete$y[j - 1]
            coordinates_complete$Central_area[j] <- coordinates_complete$Central_area[j - 1]
            coordinates_complete$Intermediate_area[j] <- coordinates_complete$Intermediate_area[j - 1]
            coordinates_complete$Refuge_area_1[j] <- coordinates_complete$Refuge_area_1[j - 1]
            coordinates_complete$Refuge_area_2[j] <- coordinates_complete$Refuge_area_2[j - 1]
            coordinates_complete$Initial_refuge[j] <-  coordinates_complete$Initial_refuge[j - 1]
            coordinates_complete$Distant_wall[j] <- coordinates_complete$Distant_wall[j - 1]
            coordinates_complete$Lateral_1[j] <- coordinates_complete$Lateral_1[j - 1]
            coordinates_complete$Lateral_2[j] <- coordinates_complete$Lateral_2[j - 1]
            coordinates_complete$Arena_area[j] <- coordinates_complete$Arena_area[j - 1]
            coordinates_complete$Area[j] <- coordinates_complete$Area[j - 1]
            coordinates_complete$dist_hole_0[j] <- coordinates_complete$dist_hole_0[j - 1]
            coordinates_complete$dist_hole_1[j] <- coordinates_complete$dist_hole_1[j - 1]
            coordinates_complete$dist_hole_2[j] <- coordinates_complete$dist_hole_2[j - 1]
            coordinates_complete$dist_hole_3[j] <- coordinates_complete$dist_hole_3[j - 1]
            coordinates_complete$dist_hole_4[j] <- coordinates_complete$dist_hole_4[j - 1]
            coordinates_complete$dist_min[j] <- 0
            coordinates_complete$hole_dist_min[j] <- coordinates_complete$hole_dist_min[j - 1]
            coordinates_complete$dist_traveled[j] <- 0
            coordinates_complete$behaviour[j] <- "hidden"
            
          }
          
        }
        
      }
      
      # It might happen that the snake is being recorded for more than 1h, and we want to know their behaviour in 1h-tests. It's for this reason that we are going to cut all videos to 1h (3600 sec). If the video lasts less than 1h, for the moment we are not going to do anything.
      if(nrow(coordinates_complete) > 3600){
        
        coordinates_complete <- coordinates_complete[1:3600, ]
        
      } else {
        
      }
      
      Snake_array[1:nrow(coordinates_complete), , i] <- as.matrix(coordinates_complete[ , ])
      
      assign("Snake_array", Snake_array, envir=globalenv())
      
      paste("Snake_array", i)
      
    }
    
  }
  
}

# Set WD to save 3D array
setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking")

# Save 3D array
save(Snake_array, file="Snake_array.Rda")

# Load snake 3D array
load("Snake_array.Rda")
```

`Arena_array`

```{r arena array}
#| warning: false
#| eval: false
# Arena array
# Variable names for each column
variable_names <- c("t", "x", "y")

# We create empty data frame that is going to store: x axis = time step, y = variables, z = individuals
Arena_array <- array(NA, dim = c(13, length(variable_names), nrow(Snake_data_tracked)))

# We set the column names for the variables across the array
dimnames(Arena_array)[[2]] <- variable_names

# We set the ID code for all tracked snakes across the array
dimnames(Arena_array)[[3]] <- Snake_data_tracked$Snake_ID

# Arena array
# THIS TAKES A LOT OF TIME. DO NOT RUN AND LOAD DIRECTLY FROM WD
for(i in 1:nrow(Snake_data_tracked)){

  # If it is the first trial (no repeatability)...
  if(Snake_data_tracked$Repeatability[i] == "N"){
    
    # ... set the working directory in the correct folder...
    setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking/Excels tracking")
    
    # ... and if it didn't get out of the original refuge, fill the correspondent data layer with 0. Otherwise...
    if(Snake_data_tracked$Time_head_out_sec_raw[i] == "NA"){
      
      # There is no arena drown for snakes that didn't get out of the refuge, so we have to draw them manually
      arena <- as.data.frame(matrix(NA, nrow = 13, ncol = 3))
      
      colnames(arena) <- c("t", "x", "y")
      
      arena$t <- c(seq(0, 12, 1))
      
      arena$x <- c(1, 1, -1, -1, 0, 0.73, 0.73, -0.73, -0.73, 0.73, 0.73, -0.73, -0.73)
      
      arena$y <- c(0, 1.5, 1.5, 0, 0, 0.43, 1.06, 1.06, 0.43, 0, 1.5, 1.5, 0)
      
      Arena_array[1:nrow(arena), , i] <- as.matrix(arena[ , ])
      
      assign("Arena_array", Arena_array, envir=globalenv())
      
      paste("Arena_array", i)
      
      
      # ... extract data from its tracked Excel file.
    } else if(Snake_data_tracked$Time_head_out_sec_raw[i] != "NA"){
      
      # Select arena limits and refuges from specific individual
      name_arena <- paste0(Snake_data_tracked$Snake_ID[i], "_arena.xlsx")
      
      arena <- read_excel(name_arena)
      
      Arena_array[1:nrow(arena), , i] <- as.matrix(arena[ , ])
      
      assign("Arena_array", Arena_array, envir=globalenv())
      
      paste("Arena_array", i)
      
    }
    
    # If it is a snake used for repeatability, ...
  } else if(Snake_data_tracked$Repeatability[i] == "Y"){
    
    # ... set the correct wd ...
    setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking/Excels tracking/Replicates")
    
    # ... and check if the snake did get out of the initial refuge or not. If it didn't, fill with 0.
    if(Snake_data_tracked$Time_head_out_sec_raw[i] == "NA"){
      
      # There is no arena drown for snakes that didn't get out of the refuge, so we have to draw them manually
      arena <- as.data.frame(matrix(NA, nrow = 13, ncol = 3))
      
      colnames(arena) <- c("t", "x", "y")
      
      arena$t <- c(seq(0, 12, 1))
      
      arena$x <- c(1, 1, -1, -1, 0, 0.73, 0.73, -0.73, -0.73, 0.73, 0.73, -0.73, -0.73)
      
      arena$y <- c(0, 1.5, 1.5, 0, 0, 0.43, 1.06, 1.06, 0.43, 0, 1.5, 1.5, 0)
      
      Arena_array[1:nrow(arena), , i] <- as.matrix(arena[ , ])
      
      assign("Arena_array", Arena_array, envir=globalenv())
      
      paste("Arena_array", i)
      
      
      # If the snake did get out of the refuge, extract data from excel
    } else if(Snake_data_tracked$Time_head_out_sec_raw[i] != "NA"){
      
      # Select arena limits and refuges from specific individual
      name_arena <- paste0(Snake_data_tracked$Snake_ID[i], "_arena.xlsx")
      
      arena <- read_excel(name_arena)
      
      Arena_array[1:nrow(arena), , i] <- as.matrix(arena[ , ])
      
      assign("Arena_array", Arena_array, envir=globalenv())
      
      paste("Arena_array", i)
      
    }
    
  }
  
  
  
}

# Set WD to save 3D array
setwd("C:/Users/marc9/Desktop/Marc/CREAF/Snake tracking")

# Save 3D array
save(Arena_array, file="Arena_array.Rda")

# Load arena 3D array
load("Arena_array.Rda")
```

## Summarized data

No need to run if `summary_tracking` and `summary_tracking_complete` are already loaded.

`summary_tracking`

```{r summarized data}
#| warning: false
#| eval: false
variables_track <- c("ID", "time_central_area", "time_intermediate_area", "time_refuge_area_1", "time_refuge_area_2", "time_initial_refuge", "time_distant_wall", "time_lateral_1", "time_lateral_2", "time_hidden_bf_head_out", "total_time_hidden", "time_entrance", "time_hole_1", "time_hole_2", "time_hole_3", "time_hole_4", "mean_dist_refuge", "mean_dist_closest_refuge", "beh_moving", "beh_exploring_wall", "beh_immobile_exposed", "beh_head-out", "beh_hiding-head-out", "beh_head-out_after_body_out", "avg_speed", "avg_speed_exploring", "total_dist_travelled", "dist_travelled_exploring", "dist_trav_5min_body_out", "dist_trav_10min_body_out", "dist_trav_15min_body_out", "dist_trav_20min_body_out", "dist_trav_25min_body_out", "dist_trav_30min_body_out")

# We create the data frame that is going to store all the data
summary_tracking <- data.frame(matrix(ncol = length(variables_track), nrow = nrow(Snake_data_tracked)))

colnames(summary_tracking) <- variables_track

# We generate the final data frame based on a fixed number of seconds
generate_summary_tracking <- function(time_limit){
  
  for(i in 1:nrow(Snake_data_tracked)){
    
    if(Snake_data_tracked$Time_head_out_sec_raw[i] != "NA" & as.numeric(Snake_data_tracked$Time_head_out_sec_raw[i]) < time_limit){
      
      # Extract snake info from array
      coordinates <- as.data.frame(Snake_array[1:time_limit, , i])
      
      # Extract arena info from array
      arena <- as.data.frame(Arena_array[, , i])
      
      # How much time does the snake hide before first head-out
      sec_bf_head_out <- as.numeric(Snake_data_tracked$Time_head_out_sec_raw[i])
      
      # for those individuals that do head-out, but not body-out
      if(Snake_data_tracked$Time_body_out_sec_raw[i] != "NA"){
        
        sec_bf_body_out <- as.numeric(Snake_data_tracked$Time_body_out_sec_raw[i])
        
      } else  {
        
        sec_bf_body_out <- time_limit
        
      }
      
      
      
      # Let's compile all the info we have
      
      tracking_ind <- summary_tracking[1, ]
      
      tracking_ind[1, ] <- NA
      
      # ID
      tracking_ind$ID <- Snake_data_tracked$Snake_ID[i]
      
      # Area use
      Time_central_area <- 0
      
      Time_intermediate_area <- 0
      
      Time_refuge_area_1 <- 0
      
      Time_refuge_area_2 <- 0
      
      Time_initial_refuge <- 0
      
      Time_distant_wall <- 0
      
      Time_lateral_area_1 <- 0
      
      Time_lateral_area_2 <- 0
      
      Time_hidden <- 0
      
      # How much time does the snake spend in each area
      for(k in 1:nrow(coordinates)){
        
        if(coordinates$behaviour[k] == "hidden"){
          
          Time_hidden <- Time_hidden + 1
          
        } else {
          
          if (coordinates$Area[k] == "Central area") {
            
            Time_central_area <- Time_central_area + 1
            
          } else if (coordinates$Area[k] == "Intermediate area") {
            
            Time_intermediate_area <- Time_intermediate_area + 1
            
          } else if (coordinates$Area[k] == "Refuge area 1"){
            
            Time_refuge_area_1 <- Time_refuge_area_1 + 1
            
          } else if (coordinates$Area[k] == "Refuge area 2"){
            
            Time_refuge_area_2 <- Time_refuge_area_2 + 1
            
          } else if (coordinates$Area[k] == "Initial refuge"){
            
            Time_initial_refuge <- Time_initial_refuge + 1
            
          } else if (coordinates$Area[k] == "Distant wall"){
            
            Time_distant_wall <- Time_distant_wall + 1
            
          } else if (coordinates$Area[k] == "Lateral 1"){
            
            Time_lateral_area_1 <- Time_lateral_area_1 + 1
            
          } else if (coordinates$Area[k] == "Lateral 2"){
            
            Time_lateral_area_2 <- Time_lateral_area_2 + 1
            
          }
          
        }
        
      }
      
      tracking_ind$time_central_area <- Time_central_area
      
      tracking_ind$time_intermediate_area <- Time_intermediate_area
      
      tracking_ind$time_refuge_area_1 <- Time_refuge_area_1
      
      tracking_ind$time_refuge_area_2 <- Time_refuge_area_2
      
      tracking_ind$time_initial_refuge <- Time_initial_refuge
      
      tracking_ind$time_distant_wall <- Time_distant_wall
      
      tracking_ind$time_lateral_1 <- Time_lateral_area_1
      
      tracking_ind$time_lateral_2 <- Time_lateral_area_2
      
      tracking_ind$time_hidden_bf_head_out <- as.numeric(sec_bf_head_out)
      
      tracking_ind$total_time_hidden <- Time_hidden
      
      # Proximity to refuges
      no_hidden <- coordinates[!grepl("hidden", coordinates$behaviour), ]
      
      hole_proximity <- table(no_hidden$hole_dist_min)
      
      tracking_ind$time_entrance <- hole_proximity[1]
      
      tracking_ind$time_hole_1 <- hole_proximity[2]
      
      tracking_ind$time_hole_2 <- hole_proximity[3]
      
      tracking_ind$time_hole_3 <- hole_proximity[4]
      
      tracking_ind$time_hole_4 <- hole_proximity[5]
      
      dist_refuges <- c(no_hidden$dist_hole_0, no_hidden$dist_hole_1, no_hidden$dist_hole_2, no_hidden$dist_hole_3, no_hidden$dist_hole_4)
      
      tracking_ind$mean_dist_refuge <- mean(as.numeric(dist_refuges))
      
      tracking_ind$mean_dist_closest_refuge <- mean(as.numeric(no_hidden$dist_min))
      
      # Behaviour
      moving <- 0
      
      exploring_wall <- 0
      
      head_out <- 0
      
      hidden_head_out <- 0
      
      immobile <- 0
      
      hidden <- 0
      
      
      for(k in 1:nrow(coordinates)){
        
        if(coordinates$behaviour[k] == "moving"){
          
          moving <- moving + 1
          
        } else if(coordinates$behaviour[k] == "exploring wall"){
          
          exploring_wall <- exploring_wall + 1
          
        } else if(coordinates$behaviour[k] == "head-out"){
          
          head_out <- head_out + 1
          
        } else if(coordinates$behaviour[k] == "hiding / head-out"){
          
          hidden_head_out <- hidden_head_out + 1
          
        } else if(coordinates$behaviour[k] == "immobile exposed"){
          
          immobile <- immobile + 1
          
        } else if(coordinates$behaviour[k] == "hidden"){
          
          hidden <- hidden + 1
          
        }
        
        
      }
      
      # To find how many seconds the snake does head-out after exiting the initial refuge
      if(sec_bf_body_out != time_limit){
        
        coordinates_body_out <- coordinates[sec_bf_body_out:nrow(coordinates), ]
        
        head_out_after_body_out <- 0
        
        for(j in 1:nrow(coordinates_body_out)){
          
          if(coordinates_body_out$behaviour[j] == "head-out"){
            
            head_out_after_body_out <- head_out_after_body_out + 1
            
          } else {
            
          }
          
        }
        
      } else {
        
        coordinates_body_out <- coordinates[1, ]
          
        head_out_after_body_out <- 0
        
      }

      
      
      
      tracking_ind$beh_moving <- moving
      
      tracking_ind$beh_exploring_wall <- exploring_wall
      
      tracking_ind$beh_immobile_exposed <- immobile
      
      tracking_ind$`beh_head-out` <- head_out
      
      tracking_ind$`beh_hiding-head-out` <- hidden_head_out
      
      tracking_ind$`beh_head-out_after_body_out` <- head_out_after_body_out
      
      tracking_ind$avg_speed <- mean(as.numeric(coordinates$dist_traveled))
      
      tracking_ind$total_dist_travelled <- sum(as.numeric(coordinates$dist_traveled))
      
      # Distance travelled when moving in the floor, not wall exploration
      moving <- coordinates[coordinates$behaviour == "moving", ]
      
      tracking_ind$avg_speed_exploring <- mean(as.numeric(moving$dist_traveled))
      
      tracking_ind$dist_travelled_exploring <- sum(as.numeric(moving$dist_traveled))
      
      # if the snake is out for more than 5 min
      if(nrow(coordinates_body_out) >= 300){
        
        # distance travelled in those 5 min
        tracking_ind$dist_trav_5min_body_out <- sum(as.numeric(coordinates_body_out$dist_traveled[1:300]))
        
        # Then, if the snake has been out for 10 min or more
        if(nrow(coordinates_body_out) >= 600){
          
          # distance travelled in those 10 min
          tracking_ind$dist_trav_10min_body_out <- sum(as.numeric(coordinates_body_out$dist_traveled[1:600]))
          
          # and if it has been outside for more than 15 min
          if(nrow(coordinates_body_out) >= 900){
            
            # distance travelled in those 15 min
            tracking_ind$dist_trav_15min_body_out <- sum(as.numeric(coordinates_body_out$dist_traveled[1:900]))
            
            if(nrow(coordinates_body_out) >= 1200){
              
              # distance travelled in those 20 min
              tracking_ind$dist_trav_20min_body_out <- sum(as.numeric(coordinates_body_out$dist_traveled[1:1200]))
              
              if(nrow(coordinates_body_out) >= 1500){
                
                # distance travelled in those 25 min
                tracking_ind$dist_trav_25min_body_out <- sum(as.numeric(coordinates_body_out$dist_traveled[1:1500]))
                
                if(nrow(coordinates_body_out) >= 1800){
                  
                  # distance travelled in those 25 min
                  tracking_ind$dist_trav_30min_body_out <- sum(as.numeric(coordinates_body_out$dist_traveled[1:1800]))
                  
                }
                
                
              }
              
              
            }
            
          }
          
        }
        
      } else {
        
        # do nothing
        
      }
      
      
      # Merge ind data to global database
      
      summary_tracking[i, ] <- tracking_ind
      
      summary_tracking[i, ][summary_tracking[i, ] == "NaN"] <- NA
      
      summary_tracking[i, ][is.na(summary_tracking[i, ])] <- 0
      
      assign("summary_tracking", summary_tracking, envir=globalenv())
      
      
      
    } else {
      
      
      tracking_ind <- summary_tracking[1, ]
      
      tracking_ind[1, ] <- NA
      
      # ID
      tracking_ind$ID <- Snake_data_tracked$Snake_ID[i]
      
      # Area use
      tracking_ind$time_central_area <- 0
      
      tracking_ind$time_intermediate_area <- 0
      
      tracking_ind$time_refuge_area_1 <- 0
      
      tracking_ind$time_refuge_area_2 <- 0
      
      tracking_ind$time_initial_refuge <- 0
      
      tracking_ind$time_distant_wall <- 0
      
      tracking_ind$time_lateral_1 <- 0
      
      tracking_ind$time_lateral_2 <- 0
      
      tracking_ind$time_hidden_bf_head_out <- time_limit
      
      tracking_ind$total_time_hidden <- time_limit
      
      # Proximity to refuges
      tracking_ind$time_entrance <- 0
      
      tracking_ind$time_hole_1 <- 0
      
      tracking_ind$time_hole_2 <- 0
      
      tracking_ind$time_hole_3 <- 0
      
      tracking_ind$time_hole_4 <- 0
      
      tracking_ind$mean_dist_refuge <- 0
      
      tracking_ind$mean_dist_closest_refuge <- 0
      
      # Behaviour
      tracking_ind$beh_moving <- 0
      
      tracking_ind$beh_exploring_wall <- 0
      
      tracking_ind$beh_immobile_exposed <- 0
      
      tracking_ind$`beh_head-out` <- 0
      
      tracking_ind$`beh_hiding-head-out` <- 0
      
      tracking_ind$`beh_head-out_after_body_out` <- 0
      
      tracking_ind$avg_speed <- 0
      
      tracking_ind$avg_speed_exploring <- 0
      
      tracking_ind$total_dist_travelled <- 0
      
      tracking_ind$dist_travelled_exploring <- 0
      
      tracking_ind$dist_trav_5min_body_out <- 0
      
      tracking_ind$dist_trav_10min_body_out <- 0
      
      tracking_ind$dist_trav_15min_body_out <- 0
      
      tracking_ind$dist_trav_20min_body_out <- 0
      
      tracking_ind$dist_trav_25min_body_out <- 0
      
      tracking_ind$dist_trav_30min_body_out <- 0
      
      
      # Merge ind data to global database
      
      summary_tracking[i, ] <- tracking_ind
      
      assign("summary_tracking", summary_tracking, envir=globalenv())
      
    }
    
  }
  
}

generate_summary_tracking(3000)

# Save file in correspondant folder
write_xlsx(summary_tracking, "C:\\Users\\marc9\\Desktop\\Marc\\CREAF\\Snake tracking\\summary_tracking.xlsx")
```

`summary_tracking_complete`

```{r summarized data merging}
#| warning: false
#| eval: false
time_limit <- 3000

# Select columns of interest
Snake_data_tracked_final <- Snake_data_tracked[, c("Snake_ID", "Trap_ID", "Locality", "X", "Y", "Inv_situation", "Capture_day", "Test_day", "Acc_time", "Hour manip.", "Hour accl.", "Hour test", "Time_head_out_sec_raw", "Time_body_out_sec_raw", "Temperature", "R_H", "Weather", "Arena", "Official", "Repeatability", "Tracking", "Year_invasion", "Distance_Noahs", "SVL", "Real_Weight", "Condition", "Sex", "N_eggs", "Hard_Drive")]

# Transform variable into numeric
Snake_data_tracked_final$Time_head_out_sec_raw <- as.numeric(Snake_data_tracked_final$Time_head_out_sec_raw)

Snake_data_tracked_final$Time_body_out_sec_raw <- as.numeric(Snake_data_tracked_final$Time_body_out_sec_raw)

# If snake didn't get out, the number of seconds is 3000
Snake_data_tracked_final <- Snake_data_tracked_final %>% mutate(Time_head_out_sec_raw = ifelse(is.na(Time_head_out_sec_raw), time_limit, Time_head_out_sec_raw))

Snake_data_tracked_final <- Snake_data_tracked_final %>% mutate(Time_body_out_sec_raw = ifelse(is.na(Time_body_out_sec_raw), time_limit, Time_body_out_sec_raw))

# If head-out/Body-out time is larger than 3000 sec, cut it at 3000 
for(i in 1:nrow(Snake_data_tracked_final)){
  
  if(Snake_data_tracked_final$Time_head_out_sec_raw[i] > time_limit){
    
    Snake_data_tracked_final$Time_head_out_sec_raw[i] <- time_limit
    
  } else {
    
  }
  
  if(Snake_data_tracked_final$Time_body_out_sec_raw[i] > time_limit){
    
    Snake_data_tracked_final$Time_body_out_sec_raw[i] <- time_limit
    
  } else {
    
  }
  
}


# Change type of variable for some variables
Snake_data_tracked_final$Locality <- as.factor(Snake_data_tracked_final$Locality)
Snake_data_tracked_final$Inv_situation <- as.factor(Snake_data_tracked_final$Inv_situation)
Snake_data_tracked_final$Capture_day <- as.factor(Snake_data_tracked_final$Capture_day)
Snake_data_tracked_final$Test_day <- as.factor(Snake_data_tracked_final$Test_day)
Snake_data_tracked_final$Acc_time <- as.factor(Snake_data_tracked_final$Acc_time)
Snake_data_tracked_final$Arena <- as.factor(Snake_data_tracked_final$Arena)
Snake_data_tracked_final$Official <- as.factor(Snake_data_tracked_final$Official)
Snake_data_tracked_final$Repeatability <- as.factor(Snake_data_tracked_final$Repeatability)
Snake_data_tracked_final$Tracking <- as.factor(Snake_data_tracked_final$Tracking)


summary_tracking_complete <- cbind(Snake_data_tracked_final, summary_tracking)

summary_tracking_complete <- add_column(summary_tracking_complete, Visual_exp = summary_tracking_complete$Time_body_out_sec_raw - summary_tracking_complete$Time_head_out_sec_raw, .after = summary_tracking_complete$Time_body_out_sec_raw)

# Save file as CSV in order to plot it in QGIS and extract year of invasion, distance to Noah's garden, ...
write.csv2(summary_tracking_complete, "C:\\Users\\marc9\\Desktop\\Marc\\CREAF\\Snake tracking\\summary_tracking_complete.csv", row.names = FALSE)

write_xlsx(summary_tracking_complete, "C:\\Users\\marc9\\Desktop\\Marc\\CREAF\\Snake tracking\\summary_tracking_complete.xlsx")

```

# Individual plots

This part is dedicated to the visualisation of the movement of each individual inside the arena. There are also plots dedicated to showing the proportion of time taken by each snake to explore the different areas of the arena, the behaviour expressed and the speed of movement at each instant.

For the visualization of these plots, you must choose the ID of the individual, if it is the first behavioural test that it did (N) or if it is one of the individuals that repeated the experiments to study the repeatability of the behaviour (Y), and the time in seconds that you want to analyse (maximum 3000 seconds).

## Extraction functions

```{r individual funcitons}
#| warning: false
ID <- 202308
repeatability <- "N"
time_limit <- 3000

data_visualization <- function(ID, repeatability, time_limit){
  
  # Which snake are we going to visualize
  selected_ind <- which(Snake_data_tracked$Snake_ID == ID & Snake_data_tracked$Repeatability == repeatability)
  
  # Extract snake info from array
  coordinates <- as.data.frame(Snake_array[1:time_limit, , selected_ind])
  
  # Extract arena info from array
  arena <- as.data.frame(Arena_array[, , selected_ind])
  
  # Let's start drawing
  plot(c(-1.1, 1.1), c(-0.1, 1.6), col = "white", xlab = "X", ylab = "Y") 
  
  # Lateral refuge area 1
  polygon(x = c(arena[1, 2], arena[2, 2], arena[2, 2] - 0.25, arena[1, 2] - 0.25),
          y = c(arena[1, 3], arena[2, 3], arena[2, 3], arena[1, 3]),
          col = viridis(6)[6], border = viridis(6)[6]) 
  
  # Lateral refuge area 2
  polygon(x = c(arena[3, 2], arena[4, 2], arena[4, 2] + 0.25, arena[3, 2] + 0.25),
          y = c(arena[3, 3], arena[4, 3], arena[4, 3], arena[3, 3]),
          col = viridis(6)[6], border = viridis(6)[6])
  
  # Distant wall area
  polygon(x = c(arena[2, 2] - 0.25, arena[2, 2] - 0.25, arena[3, 2] + 0.25, arena[3, 2] + 0.25),
          y = c(arena[2, 3] - 0.25, arena[2, 3], arena[3, 3], arena[3, 3] - 0.25),
          col = viridis(6)[5], border = viridis(6)[5]) 
  
  # Initial refuge area
  polygon(x =c(arena[1, 2] - 0.25, arena[1, 2] -0.25, arena[4, 2] + 0.25, arena[4, 2] + 0.25),
          y = c(arena[1, 3], arena[1, 3] + 0.25, arena[4, 3] + 0.25, arena[4, 3]),
          col = viridis(6)[4], border = viridis(6)[4]) 
  
  # Refuge area 1
  polygon(x = c(arena[1, 2]- 0.25, arena[2, 2] - 0.25, arena[2, 2] - 0.5, arena[1, 2] - 0.5),
          y = c(arena[1, 3] + 0.25, arena[2, 3] - 0.25, arena[2, 3] - 0.25, arena[1, 3] + 0.25),
          col = viridis(6)[3], border = viridis(6)[3]) 
  
  # Refuge area 2
  polygon(x = c(arena[4, 2] + 0.5, arena[3, 2] + 0.5, arena[3, 2] + 0.25, arena[4, 2] + 0.25),
          y = c(arena[4, 3] + 0.25, arena[3, 3] - 0.25, arena[3, 3] - 0.25, arena[4, 3] + 0.25),
          col = viridis(6)[3], border = viridis(6)[3])
  
  # Intermediate area
  polygon(x = c(arena[1, 2]- 0.5, arena[2, 2] - 0.5, arena[3, 2] + 0.5, arena[4, 2] + 0.5),
          y = c(arena[1, 3] + 0.25, arena[2, 3] - 0.25, arena[3, 3] - 0.25, arena[4, 3] + 0.25),
          col = viridis(6)[2], border = viridis(6)[2]) 
  
  # Central area
  polygon(x = c(arena[1, 2]- 0.75, arena[2, 2] - 0.75, arena[3, 2] + 0.75, arena[4, 2] + 0.75),
          y = c(arena[1, 3] + 0.5, arena[2, 3] - 0.5, arena[3, 3] - 0.5, arena[4, 3] + 0.5),
          col = viridis(6)[1], border = viridis(6)[1]) 
  
  
  # Border real arena
  polygon(x = c(arena[1, 2], arena[2, 2], arena[3, 2], arena[4, 2]),
          y = c(arena[1, 3], arena[2, 3], arena[3, 3], arena[4, 3]),  
          lwd = 2) 
  
  # Refuge coordinates
  refuges <- data.frame(cbind(c(arena[5, 2], arena[6, 2], arena[7, 2], arena[8, 2], arena[9, 2]),
                              c(arena[5, 3], arena[6, 3], arena[7, 3], arena[8, 3], arena[9, 3])))
  
  # Plot refuges
  points(refuges, pch = 19, cex = 4, col = "white")
  
  # Plot snakee trail
  points(coordinates$x, coordinates$y, type ="l", col = "black", lwd = 2)
  
  # In case you want to draw lateral refuge limits 
  # lines(c(arena[10, 2], arena[11, 2]), c(arena[10, 3], arena[11, 3]))
  
  # lines(c(arena[12, 2], arena[13, 2]), c(arena[12, 3], arena[13, 3]))
  
  # Exit from refuge (Last frame you were missing, and now I have detected you)
  for(i in 2:nrow(coordinates)){
    
    if(coordinates$behaviour[i - 1] == "hidden" && coordinates$behaviour[i] != "hidden"){
      
      points(coordinates$x[i], coordinates$y[i], 
             cex = 1, pch = 24, bg = "white", col = "black", lwd = 2)
      
    } else {
      
      
    }
    
  }
  
  
  # Enter to refuge (Now I see you, but next frame I don't)
  for(i in 1:(nrow(coordinates) - 1)){
    
    if(coordinates$behaviour[i] != "hidden" && coordinates$behaviour[i + 1] == "hidden"){
      
      points(coordinates$x[i], coordinates$y[i], 
             cex = 1, pch = 25, bg = "white", col = "black", lwd = 2)
      
    } else {
      
      
    }
    
  }
  
}

density_function <- function(ID, repeatability, time_limit){
  
  # Which snake are we going to visualize
  selected_ind <- which(Snake_data_tracked$Snake_ID == ID & Snake_data_tracked$Repeatability == repeatability)
  
  # Extract snake info from array
  coordinates <- as.data.frame(Snake_array[1:time_limit, , selected_ind])
  
  no_hidden <- coordinates[!grepl("hidden", coordinates$behaviour), ]
  
  coordinates <- no_hidden
  
  ggplot(coordinates, aes(x=as.numeric(x), y=as.numeric(y))) +
    geom_bin2d(bins = 20) +
    scale_fill_continuous(type = "viridis") +
    labs(x = "x", y = "y") +
    theme_bw()
  
  # ggplot(coordinates, aes(x=x, y=y) ) +
  # stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")
  
}

graph_area <- function(ID, repeatability, time_limit){
  
  # Which snake are we going to visualize
  selected_ind <- which(Snake_data_tracked$Snake_ID == ID & Snake_data_tracked$Repeatability == repeatability)
  
  # Extract snake info from array
  coordinates <- as.data.frame(Snake_array[1:time_limit, , selected_ind])
  
  # How many seconds the snake has the snake been in each region
  
  Time_central_area <- 0
  
  Time_intermediate_area <- 0
  
  Time_refuge_area_1 <- 0
  
  Time_refuge_area_2 <- 0
  
  Time_initial_refuge <- 0
  
  Time_distant_wall <- 0
  
  Time_lateral_area_1 <- 0
  
  Time_lateral_area_2 <- 0
  
  Time_hidden <- 0
  
  # How much time does the snake spend in each area
  for(i in 1:nrow(coordinates)){
    
    if(coordinates$behaviour[i] == "hidden"){
      
      Time_hidden <- Time_hidden + 1
      
    } else {
      
      if (coordinates$Area[i] == "Central area") {
        
        Time_central_area <- Time_central_area + 1
        
      } else if (coordinates$Area[i] == "Intermediate area") {
        
        Time_intermediate_area <- Time_intermediate_area + 1
        
      } else if (coordinates$Area[i] == "Refuge area 1"){
        
        Time_refuge_area_1 <- Time_refuge_area_1 + 1
        
      } else if (coordinates$Area[i] == "Refuge area 2"){
        
        Time_refuge_area_2 <- Time_refuge_area_2 + 1
        
      } else if (coordinates$Area[i] == "Initial refuge"){
        
        Time_initial_refuge <- Time_initial_refuge + 1
        
      } else if (coordinates$Area[i] == "Distant wall"){
        
        Time_distant_wall <- Time_distant_wall + 1
        
      } else if (coordinates$Area[i] == "Lateral 1"){
        
        Time_lateral_area_1 <- Time_lateral_area_1 + 1
        
      } else if (coordinates$Area[i] == "Lateral 2"){
        
        Time_lateral_area_2 <- Time_lateral_area_2 + 1
        
      }
      
    }
    
  }
  
  
  # Vector with number of seconds
  total_use  <- c(Time_hidden, Time_lateral_area_1, Time_lateral_area_2, Time_distant_wall, Time_initial_refuge, Time_refuge_area_1, Time_refuge_area_2, Time_intermediate_area, Time_central_area)
  # Vector with percentage 
  percent_total_use <- round(total_use/sum(total_use)*100, 1)
  
  percent_total_use <- as.data.frame(percent_total_use)
  
  area_names <- c("Hidden", "Lateral refuge 1", "Lateral refuge 2", "Distant wall", "Initial refuge",  "Refuge 1", "Refuge 2", "Intermediate",  "Central")
  
  ggplot(data = percent_total_use, aes(x = area_names, y = percent_total_use)) +
    geom_bar(stat="identity", fill = c("grey50", viridis(6)[6], viridis(6)[6], viridis(6)[5], viridis(6)[4], viridis(6)[3], viridis(6)[3], viridis(6)[2], viridis(6)[1])) + 
    geom_text(aes(label = percent_total_use), vjust = -0.3, size = 3.5) +
    scale_x_discrete(limits = area_names) + 
    labs(title= paste0("Percentage of time spent by snake ", ID , " in each area (", time_limit, " sec)"), 
         x="Area", y = "Percentage (%)") + 
    theme_classic()
  
}

graph_hole <- function(ID, repeatability, time_limit){
  
  # Which snake are we going to visualize
  selected_ind <- which(Snake_data_tracked$Snake_ID == ID & Snake_data_tracked$Repeatability == repeatability)
  
  # Extract snake info from array
  coordinates <- as.data.frame(Snake_array[1:time_limit, , selected_ind])
  
  # How much time does the snake hide before first head-out
  sec_bf_head_out <- as.numeric(Snake_data_tracked$Time_head_out_sec_raw[selected_ind])
  
  # We are not interested in the proximity to any hole when the snake is inside a hole, hidden, so we delete the records when the snake is hidden
  no_hidden <- coordinates[!grepl("hidden", coordinates$behaviour), ]
  
  coordinates <- no_hidden
  
  # Which is the closest hole
  hole_proximity <- table(coordinates$hole_dist_min)
  
  percent_hole_proximity <- round(hole_proximity/sum(hole_proximity)*100, 1)
  
  percent_hole_proximity <- as.data.frame(percent_hole_proximity)
  
  percent_hole_proximity$Var1 <- c("Entrance", "Hole 1", "Hole 2", "Hole 3", "Hole 4")
  
  ggplot(data = percent_hole_proximity, aes(x = Var1, y = Freq)) +
    geom_bar(stat="identity", fill = c(viridis(5)[5], viridis(5)[4], viridis(5)[3], viridis(5)[2], viridis(5)[1])) + 
    geom_text(aes(label = Freq), vjust = -0.3, size = 3.5) +
    scale_x_discrete(limits = percent_hole_proximity$Var1) + 
    labs(title=paste0("Which is the closest refuge while the snake ", ID, " is out? (", time_limit, " sec)"), 
         x="Hole", y = "Percentage (%)") + 
    theme_classic()
  
}

graph_behaviour <- function(ID, repeatability, time_limit){
  
  # Which snake are we going to visualize
  selected_ind <- which(Snake_data_tracked$Snake_ID == ID & Snake_data_tracked$Repeatability == repeatability)
  
  # Extract snake info from array
  coordinates <- as.data.frame(Snake_array[1:time_limit, , selected_ind])
  
  # Which is the behaviour
  moving <- 0
  
  exploring_wall <- 0
  
  head_out <- 0
  
  hidden_head_out <- 0
  
  immobile <- 0
  
  hidden <- 0
  
  for(i in 1:nrow(coordinates)){
    
    if(coordinates$behaviour[i] == "moving"){
      
      moving <- moving + 1
      
    } else if(coordinates$behaviour[i] == "exploring wall"){
      
      exploring_wall <- exploring_wall + 1
      
    } else if(coordinates$behaviour[i] == "head-out"){
      
      head_out <- head_out + 1
      
    } else if(coordinates$behaviour[i] == "hiding / head-out"){
      
      hidden_head_out <- hidden_head_out + 1
      
    } else if(coordinates$behaviour[i] == "immobile exposed"){
      
      immobile <- immobile + 1
      
    } else if(coordinates$behaviour[i] == "hidden"){
      
      hidden <- hidden + 1
      
    }
    
  }
  
  behaviour  <- c(hidden, moving, exploring_wall, head_out, hidden_head_out, immobile)
  
  percent_behaviour <- round(behaviour/sum(behaviour)*100, 1)
  
  percent_behaviour <- as.data.frame(percent_behaviour)
  
  behaviour_names <- c("Hidden", "Moving", "Exploring wall", "Head-out", "Hidden / Head-out", "Immobile exposed")
  
  ggplot(data = percent_behaviour, aes(x = behaviour_names, y = percent_behaviour)) +
    geom_bar(stat="identity", fill = c("grey50", viridis(5)[5], viridis(5)[4], viridis(5)[3], viridis(5)[2], viridis(5)[1])) + 
    geom_text(aes(label = percent_behaviour), vjust = -0.3, size = 3.5) +
    scale_x_discrete(limits = behaviour_names) + 
    labs(title=paste0("Which behaviour is the snake ", ID, " exhibiting? (", time_limit, " sec)"), 
         x="Behaviour", y = "Percentage (%)") + 
    theme_classic()
  
}

speed_plot <- function(ID, repeatability, time_limit) {
  
  par(mfrow = c(2, 1))
  
  par(mar = c(4.5, 5, 0.8, 0.5))  
  
  # Which snake are we going to visualize
  selected_ind <- which(Snake_data_tracked$Snake_ID == ID & Snake_data_tracked$Repeatability == repeatability)
  
  # Extract snake info from array
  coordinates <- as.data.frame(Snake_array[1:time_limit, , selected_ind])
  
  # We are going to delete those recordings where the snake comes out from a refuge, because the speed that the snake might exhibit is not real, it comes from going from refuge A to B from "one frame to the next one"
  for(i in 2:nrow(coordinates)){
    
    if(coordinates$behaviour[i - 1] == "hidden" && coordinates$behaviour[i] != "hidden"){
      
      coordinates$dist_traveled[i] <- 0
      
    }
    
  }
  
  # We plot the results
  plot(c(0, nrow(coordinates)), c(-0.1, max(as.numeric(coordinates$dist_traveled))), 
       col = "white", 
       xlab = "Time (s)", 
       ylab = "Speed (m/s)",
       xaxt = "n", yaxt = "n",
       axes = FALSE)
  
  ticks_x <- seq(0, nrow(coordinates), by = nrow(coordinates)/6)
  
  axis(1, at = ticks_x, 
       labels = c(ticks_x[1],
                  round_any(ticks_x[2], 5),
                  round_any(ticks_x[3], 5),
                  round_any(ticks_x[4], 5),
                  round_any(ticks_x[5], 5),
                  round_any(ticks_x[6], 5),
                  round_any(ticks_x[7], 5)))
  
  ticks_y <- round(max(as.numeric(coordinates$dist_traveled)), 2)
  
  
  axis(2, at = c(0, round((ticks_y/3), 2), round(((ticks_y/3)*2), 2), ticks_y), las = 1)
  
  
  
  # We plot the behaviour underneath
  for(i in 1:nrow(coordinates)){
    
    if(coordinates$behaviour[i] == "hidden") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = "grey50", border = "grey50")
      
    } else if(coordinates$behaviour[i] == "moving") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[5], border = viridis(5)[5])
      
    } else if(coordinates$behaviour[i] == "exploring wall") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[4], border = viridis(5)[4])
      
    } else if(coordinates$behaviour[i] == "head-out") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[3], border = viridis(5)[3])
      
    } else if(coordinates$behaviour[i] == "hiding / head-out") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[2], border = viridis(5)[2])
      
    } else if(coordinates$behaviour[i] == "immobile exposed") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[1], border = viridis(5)[1])
      
    }
    
  }
  
  # We plot the speed
  points(coordinates$t, coordinates$dist_traveled, type = "l")
  
  
  # Now for only the part where the snake has exited the initial refuge
  
  # How much time does the snake hide before first head-out
  sec_bf_head_out <- as.numeric(Snake_data_tracked$Time_head_out_sec_raw[selected_ind])
  
  coordinates_2 <- coordinates[(sec_bf_head_out + 1):nrow(coordinates), ]
  
  plot(c(0, nrow(coordinates_2)), c(-0.1, max(as.numeric(coordinates_2$dist_traveled))),
       col = "white", 
       xlab = "Time (s)", 
       ylab = "Speed (m/s)",
       yaxt = "n", xaxt = "n",
       axes = FALSE
  )
  
  ticks_x <- seq(0, nrow(coordinates_2), by = nrow(coordinates_2)/6)
  
  axis(1, at = ticks_x, 
       labels = c(ticks_x[1] + sec_bf_head_out + 1,
                  round_any(ticks_x[2] + sec_bf_head_out, 5),
                  round_any(ticks_x[3] + sec_bf_head_out, 5),
                  round_any(ticks_x[4] + sec_bf_head_out, 5),
                  round_any(ticks_x[5] + sec_bf_head_out, 5),
                  round_any(ticks_x[6] + sec_bf_head_out, 5),
                  round_any(ticks_x[7] + sec_bf_head_out, 5)))
  
  
  
  
  ticks_y <- round(max(as.numeric(coordinates_2$dist_traveled)), 2)
  
  axis(2, at = c(0, round((ticks_y/3), 2), round(((ticks_y/3)*2), 2), ticks_y), las = 1)
  
  # We plot the behaviour underneath
  for(i in 1:nrow(coordinates_2)){
    
    if(coordinates_2$behaviour[i] == "hidden") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = "grey50", border = "grey50")
      
    } else if(coordinates_2$behaviour[i] == "moving") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[5], border = viridis(5)[5])
      
    } else if(coordinates_2$behaviour[i] == "exploring wall") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[4], border = viridis(5)[4])
      
    } else if(coordinates_2$behaviour[i] == "head-out") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[3], border = viridis(5)[3])
      
    } else if(coordinates_2$behaviour[i] == "hiding / head-out") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[2], border = viridis(5)[2])
      
    } else if(coordinates_2$behaviour[i] == "immobile exposed") {
      
      polygon(x = c((i - 1), (i), (i), (i - 1)),
              y = c(- 0.02, -0.02, -0.08, -0.08),
              col = viridis(5)[1], border = viridis(5)[1])
      
    }
    
  }
  
  # We plot the speed
  points(1:nrow(coordinates_2), coordinates_2$dist_traveled, type = "l")
  
  par(mfrow = c(1, 1))
  
  
}

```

## Individual movement plots

```{r individual plots}
#| warning: false
data_visualization(ID, repeatability, time_limit)

density_function(ID, repeatability, time_limit)

graph_area(ID, repeatability, time_limit)

graph_hole(ID, repeatability, time_limit)

graph_behaviour(ID, repeatability, time_limit)

speed_plot(ID, repeatability, time_limit)
```

# Repeatability

In this part we study whether head-out, body-out and visual exploration behaviour are maintained in the same range of values for the same individual, demonstrating a certain repeatability in the experiments and a certain animal personality. In addition, it also allows us to see if there is any trend between different experimental trials depending on the area the snake comes from.

## 2023 snakes only

In this section, we will look at whether there is a difference in repeatability in head-out, body-out and visual exploration behaviours between sexes, including and excluding pregnant females to see if there is an effect of these on repeatability. Repeatability between CORE and FRONT is explored in the next section, as this only includes snakes for which we know the sex and gestation stage.

```{r repeatability 2023}
#| warning: false

# Read data from 2023 snakes
repeat_ID <- summary_tracking_complete$Snake_ID[duplicated(summary_tracking_complete$Snake_ID)]

# Select ind with more than one tracking
repeat_ind <- summary_tracking_complete[summary_tracking_complete$Snake_ID %in% repeat_ID, ]

# Assign 1 to first trials
repeat_ind$repeat_trial <- 1

for (i in 1:nrow(repeat_ind)){
  
  # And if duplicated, assign 2
  if(duplicated(repeat_ind$ID)[i] == T){
    
    repeat_ind$repeat_trial[i] <- 2
    
  } else {
    
  }
  
}

# Select variables of interest
repeat_ind <- repeat_ind[, c("Snake_ID", "Inv_situation", "Capture_day", "Time_head_out_sec_raw", "Time_body_out_sec_raw", "Repeatability", "Tracking", "Year_invasion", "Distance_Noahs", "SVL", "Real_Weight", "Condition", "Sex", "N_eggs", "repeat_trial")]

# Order snakes by ID (first the first trial and second the repeatiblity test)
repeat_ind <- repeat_ind[order(repeat_ind$Snake_ID), ]

```

### Head-out

#### Including gravid females

```{r rep headout gravid 2023}
#| warning: false

# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(repeat_ind[, c("Snake_ID", "Time_head_out_sec_raw", "repeat_trial", "Sex")])

# divide by sex
icc_data_F <- icc_data[icc_data$Sex == "F", ]

icc_data_M <- icc_data[icc_data$Sex == "M", ]

# Delete sex column
icc_data_F <- icc_data_F[, -c(4)]

icc_data_M <- icc_data_M[, -c(4)]

# Put it in wide format
icc_data_wide_F <- reshape(icc_data_F, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

icc_data_wide_M <- reshape(icc_data_M, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide_F <- icc_data_wide_F[, -1]

icc_data_wide_M <- icc_data_wide_M[, -1]

# Change column names
colnames(icc_data_wide_F) <- c("First_trial", "Second_trial")

colnames(icc_data_wide_M) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide_F <- log1p(icc_data_wide_F)

icc_data_wide_M <- log1p(icc_data_wide_M)

# Consistency (ICC)
# Females
irr::icc(icc_data_wide_F, model = "oneway", type = "consistency", unit = "single")

# Males
irr::icc(icc_data_wide_M, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = repeat_ind, aes(x = Repeatability, y = log1p(Time_head_out_sec_raw))) +
  theme_bw() + 
  facet_wrap(~ Sex, labeller = labeller(Sex = 
                                          c("F" = "Female\nICC = 0.688 | p < 0.05",
                                            "M" = "Male\nICC = 0.721 | p < 0.01"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Time to head-out (s))") + 
  ggtitle("Head-out with gravid females") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# Females
anova_F <- aov(log(Time_head_out_sec_raw) ~ repeat_trial, data = icc_data_F)

summary(anova_F)

# Males
anova_M <- aov(log(Time_head_out_sec_raw) ~ repeat_trial, data = icc_data_M)

summary(anova_M)
```

#### Excluding gravid females

```{r rep headout nogravid 2023}
#| warning: false

# Data frame without gravid females
repeat_ind_no_gravid <- repeat_ind[repeat_ind$N_eggs == 0, ]

# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(repeat_ind[, c("Snake_ID", "Time_head_out_sec_raw", "repeat_trial", "Sex", "N_eggs")])

# Do not include gravid females
icc_data <- icc_data[icc_data$N_eggs == 0, ]

# divide by sex
icc_data_F <- icc_data[icc_data$Sex == "F", ]

icc_data_M <- icc_data[icc_data$Sex == "M", ]

# Delete sex and n_eggs columns
icc_data_F <- icc_data_F[, -c(4, 5)]

icc_data_M <- icc_data_M[, -c(4, 5)]

# Put it in wide format
icc_data_wide_F <- reshape(icc_data_F, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

icc_data_wide_M <- reshape(icc_data_M, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide_F <- icc_data_wide_F[, -1]

icc_data_wide_M <- icc_data_wide_M[, -1]

# Change column names
colnames(icc_data_wide_F) <- c("First_trial", "Second_trial")

colnames(icc_data_wide_M) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide_F <- log1p(icc_data_wide_F)

icc_data_wide_M <- log1p(icc_data_wide_M)

# Consistency (ICC)
# Females
irr::icc(icc_data_wide_F, model = "oneway", type = "consistency", unit = "single")

# Males
irr::icc(icc_data_wide_M, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = repeat_ind_no_gravid, aes(x = Repeatability, y = log1p(Time_head_out_sec_raw))) +
  theme_bw() + 
  facet_wrap(~ Sex, labeller = labeller(Sex = 
                                          c("F" = "Female\nICC = 0.785 | p < 0.05",
                                            "M" = "Male\nICC = 0.721 | p < 0.01"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Time to head-out (s))") + 
  ggtitle("Head-out without gravid females") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# Females
anova_F <- aov(log(Time_head_out_sec_raw) ~ repeat_trial, data = icc_data_F)

summary(anova_F)

# Males
anova_M <- aov(log(Time_head_out_sec_raw) ~ repeat_trial, data = icc_data_M)

summary(anova_M)

```

### Body-out

#### Including gravid females

```{r rep bodyout gravid 2023}
#| warning: false

# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(repeat_ind[, c("Snake_ID", "Time_body_out_sec_raw", "repeat_trial", "Sex")])

# divide by sex
icc_data_F <- icc_data[icc_data$Sex == "F", ]

icc_data_M <- icc_data[icc_data$Sex == "M", ]

# Delete sex column
icc_data_F <- icc_data_F[, -c(4)]

icc_data_M <- icc_data_M[, -c(4)]

# Put it in wide format
icc_data_wide_F <- reshape(icc_data_F, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

icc_data_wide_M <- reshape(icc_data_M, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide_F <- icc_data_wide_F[, -1]

icc_data_wide_M <- icc_data_wide_M[, -1]

# Change column names
colnames(icc_data_wide_F) <- c("First_trial", "Second_trial")

colnames(icc_data_wide_M) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide_F <- log1p(icc_data_wide_F)

icc_data_wide_M <- log1p(icc_data_wide_M)

# Consistency (ICC)
# Females
irr::icc(icc_data_wide_F, model = "oneway", type = "consistency", unit = "single")

# Males
irr::icc(icc_data_wide_M, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = repeat_ind, aes(x = Repeatability, y = log1p(Time_body_out_sec_raw))) +
  theme_bw() + 
  facet_wrap(~ Sex, labeller = labeller(Sex = 
                                          c("F" = "Female\nICC = 0.826 | p < 0.01",
                                            "M" = "Male\nICC = 0.648 | p < 0.01"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Time to body-out (s))") + 
  ggtitle("Body-out without gravid females") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# Females
anova_F <- aov(log(Time_body_out_sec_raw) ~ repeat_trial, data = icc_data_F)

summary(anova_F)

# Males
anova_M <- aov(log(Time_body_out_sec_raw) ~ repeat_trial, data = icc_data_M)

summary(anova_M)
```

#### Excluding gravid females

```{r rep bodyout nogravid 2023}
#| warning: false

# Data frame without gravid females
repeat_ind_no_gravid <- repeat_ind[repeat_ind$N_eggs == 0, ]

# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(repeat_ind[, c("Snake_ID", "Time_body_out_sec_raw", "repeat_trial", "Sex", "N_eggs")])

# Do not include gravid females
icc_data <- icc_data[icc_data$N_eggs == 0, ]

# divide by sex
icc_data_F <- icc_data[icc_data$Sex == "F", ]

icc_data_M <- icc_data[icc_data$Sex == "M", ]

# Delete sex and n_eggs columns
icc_data_F <- icc_data_F[, -c(4, 5)]

icc_data_M <- icc_data_M[, -c(4, 5)]

# Put it in wide format
icc_data_wide_F <- reshape(icc_data_F, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

icc_data_wide_M <- reshape(icc_data_M, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide_F <- icc_data_wide_F[, -1]

icc_data_wide_M <- icc_data_wide_M[, -1]

# Change column names
colnames(icc_data_wide_F) <- c("First_trial", "Second_trial")

colnames(icc_data_wide_M) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide_F <- log1p(icc_data_wide_F)

icc_data_wide_M <- log1p(icc_data_wide_M)

# Consistency (ICC)
# Females
irr::icc(icc_data_wide_F, model = "oneway", type = "consistency", unit = "single")

# Males
irr::icc(icc_data_wide_M, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = repeat_ind_no_gravid, aes(x = Repeatability, y = log1p(Time_body_out_sec_raw))) +
  theme_bw() + 
  facet_wrap(~ Sex, labeller = labeller(Sex = 
                                          c("F" = "Female\nICC = 0.807 | p < 0.05",
                                            "M" = "Male\nICC = 0.648 | p < 0.01"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Time to body-out (s))") + 
  ggtitle("Body-out without gravid females") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# Females
anova_F <- aov(log(Time_body_out_sec_raw) ~ repeat_trial, data = icc_data_F)

summary(anova_F)

# Males
anova_M <- aov(log(Time_body_out_sec_raw) ~ repeat_trial, data = icc_data_M)

summary(anova_M)
```

### Visual exploration

#### Including gravid females

```{r rep visexp gravid 2023}
#| warning: false

# Only those ind that got out of the initial refuge on time
summary_tracking_complete_outside <- summary_tracking_complete[summary_tracking_complete$Time_body_out_sec_raw != time_limit, ]

# select repeated individuals within the ones that got out
repeat_ID_outside <- summary_tracking_complete_outside$Snake_ID[duplicated(summary_tracking_complete_outside$Snake_ID)]

# select individuals
repeat_ind_outside <- summary_tracking_complete_outside[summary_tracking_complete_outside$Snake_ID %in% repeat_ID_outside, ]

# assign 1 to first trials
repeat_ind_outside$repeat_trial <- 1

for (i in 1:nrow(repeat_ind_outside)){
  
  # And if duplicated, assign 2
  if(duplicated(repeat_ind_outside$ID)[i] == T){
    
    repeat_ind_outside$repeat_trial[i] <- 2
    
  } else {
    
  }
  
}

# Order snakes by ID (first the first trial and second the repeatiblity test)
repeat_ind_outside <- repeat_ind_outside[order(repeat_ind_outside$ID), ]

# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(repeat_ind_outside[, c("Snake_ID", "Visual_exp", "repeat_trial", "Sex")])

# divide by sex
icc_data_F <- icc_data[icc_data$Sex == "F", ]

icc_data_M <- icc_data[icc_data$Sex == "M", ]

# Delete sex column
icc_data_F <- icc_data_F[, -c(4)]

icc_data_M <- icc_data_M[, -c(4)]

# Put it in wide format
icc_data_wide_F <- reshape(icc_data_F, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

icc_data_wide_M <- reshape(icc_data_M, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide_F <- icc_data_wide_F[, -1]

icc_data_wide_M <- icc_data_wide_M[, -1]

# Change column names
colnames(icc_data_wide_F) <- c("First_trial", "Second_trial")

colnames(icc_data_wide_M) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide_F <- log1p(icc_data_wide_F)

icc_data_wide_M <- log1p(icc_data_wide_M)

# Consistency (ICC)
# Females
irr::icc(icc_data_wide_F, model = "oneway", type = "consistency", unit = "single")

# Males
irr::icc(icc_data_wide_M, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = repeat_ind_outside, aes(x = Repeatability, y = log1p(Visual_exp))) +
  theme_bw() + 
  facet_wrap(~ Sex, labeller = labeller(Sex = 
                                          c("F" = "Female\nICC = -0.0178 | p = 0.491",
                                            "M" = "Male\nICC = 0.313 | p = 0.192"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Visual exploration time (s))") + 
  ggtitle("Visual exploration with gravid females") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# Females
anova_F <- aov(log1p(Visual_exp) ~ repeat_trial, data = icc_data_F)

summary(anova_F)

# Males
anova_M <- aov(log1p(Visual_exp) ~ repeat_trial, data = icc_data_M)

summary(anova_M)
```

#### Excluding gravid females

```{r rep visexp nogravid 2023}
#| warning: false

# Only those ind that got out of the initial refuge on time
summary_tracking_complete_outside <- summary_tracking_complete[summary_tracking_complete$Time_body_out_sec_raw != time_limit, ]

# select repeated individuals within the ones that got out
repeat_ID_outside <- summary_tracking_complete_outside$Snake_ID[duplicated(summary_tracking_complete_outside$Snake_ID)]

# select individuals
repeat_ind_outside <- summary_tracking_complete_outside[summary_tracking_complete_outside$Snake_ID %in% repeat_ID_outside, ]

# Data frame without gravid females
repeat_ind_outside_no_gravid <- repeat_ind_outside[repeat_ind_outside$N_eggs == 0, ]

# assign 1 to first trials
repeat_ind_outside$repeat_trial <- 1

for (i in 1:nrow(repeat_ind_outside)){
  
  # And if duplicated, assign 2
  if(duplicated(repeat_ind_outside$ID)[i] == T){
    
    repeat_ind_outside$repeat_trial[i] <- 2
    
  } else {
    
  }
  
}

# Order snakes by ID (first the first trial and second the repeatiblity test)
repeat_ind_outside <- repeat_ind_outside[order(repeat_ind_outside$ID), ]

# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(repeat_ind_outside[, c("Snake_ID", "Visual_exp", "repeat_trial", "Sex", "N_eggs")])

# Do not include gravid females
icc_data <- icc_data[icc_data$N_eggs == 0, ]

# divide by sex
icc_data_F <- icc_data[icc_data$Sex == "F", ]

icc_data_M <- icc_data[icc_data$Sex == "M", ]

# Delete sex and n_eggs columns
icc_data_F <- icc_data_F[, -c(4, 5)]

icc_data_M <- icc_data_M[, -c(4, 5)]

# Put it in wide format
icc_data_wide_F <- reshape(icc_data_F, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

icc_data_wide_M <- reshape(icc_data_M, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide_F <- icc_data_wide_F[, -1]

icc_data_wide_M <- icc_data_wide_M[, -1]

# Change column names
colnames(icc_data_wide_F) <- c("First_trial", "Second_trial")

colnames(icc_data_wide_M) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide_F <- log1p(icc_data_wide_F)

icc_data_wide_M <- log1p(icc_data_wide_M)

# Consistency (ICC)
# Females
irr::icc(icc_data_wide_F, model = "oneway", type = "consistency", unit = "single")

# Males
irr::icc(icc_data_wide_M, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = repeat_ind_outside_no_gravid, aes(x = Repeatability, y = log1p(Visual_exp))) +
  theme_bw() + 
  facet_wrap(~ Sex, labeller = labeller(Sex = 
                                          c("F" = "Female\nICC = 0.402 | p = 0.244",
                                            "M" = "Male\nICC = 0.313 | p = 0.192"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Visual exploration time (s))") + 
  ggtitle("Visual exploration without gravid females") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# Females
anova_F <- aov(log1p(Visual_exp) ~ repeat_trial, data = icc_data_F)

summary(anova_F)

# Males
anova_M <- aov(log1p(Visual_exp) ~ repeat_trial, data = icc_data_M)

summary(anova_M)

```

## 2023 + 2024 snakes

In this section, for the moment we can only see if there are differences between CORE or FRONT, as at the moment we do not have the sex of the individuals or their pregnancy status.

```{r repeatability 2023-2024}
#| warning: false

# detect duplicated ID from 2023 data
repeat_ID <- summary_tracking_complete$Snake_ID[duplicated(summary_tracking_complete$Snake_ID)]

# Select ind with more than one tracking
repeat_ind <- summary_tracking_complete[summary_tracking_complete$Snake_ID %in% repeat_ID, ]

# assign 1 to first trials
repeat_ind$repeat_trial <- 1

for (i in 1:nrow(repeat_ind)){
  
  # And if duplicated, assign 2
  if(duplicated(repeat_ind$ID)[i] == T){
    
    repeat_ind$repeat_trial[i] <- 2
    
  } else {
    
  }
  
}

# Order snakes by ID (first the first trial and second the repeatiblity test)
repeat_ind <- repeat_ind[order(repeat_ind$ID), ]

# Reduced data frame
repeat_ind_2024_merged <- repeat_ind[, c("Snake_ID", "Inv_situation", "Capture_day", "Time_head_out_sec_raw", "Time_body_out_sec_raw", "Repeatability", "Tracking", "Year_invasion", "Distance_Noahs", "SVL", "Real_Weight", "Condition", "Sex", "N_eggs", "repeat_trial")]

# 2024 data
repeat_2024 <- Snake_data[year(Snake_data$Capture_day) == 2024, ]

# assign 1 to first trials
repeat_2024$repeat_trial <- 1

for (i in 1:nrow(repeat_2024)){
  
  # And if duplicated, assign 2
  if(duplicated(repeat_2024$Snake_ID)[i] == T){
    
    repeat_2024$repeat_trial[i] <- 2
    
  } else {
    
  }
  
}

# Reduced data frame
repeat_2024_merged_2 <- repeat_2024[, c("Snake_ID", "Inv_situation", "Capture_day", "Time_head_out_sec_raw", "Time_body_out_sec_raw", "Repeatability", "Tracking", "Year_invasion", "Distance_Noahs", "repeat_trial")]

# Date format
repeat_ind_2024_merged$Capture_day <- as_datetime(repeat_ind_2024_merged$Capture_day)

repeat_ind_2024_merged$Capture_day <- as_datetime(repeat_ind_2024_merged$Capture_day)

# Tiem to head-out / body-out as numeric 
repeat_2024_merged_2$Time_head_out_sec_raw <- as.numeric(repeat_2024_merged_2$Time_head_out_sec_raw)

repeat_2024_merged_2$Time_body_out_sec_raw <- as.numeric(repeat_2024_merged_2$Time_body_out_sec_raw)

# If snake didn't get out, the number of seconds is 3000
repeat_2024_merged_2 <- repeat_2024_merged_2 %>% mutate(Time_head_out_sec_raw = ifelse(is.na(Time_head_out_sec_raw), time_limit, Time_head_out_sec_raw))

repeat_2024_merged_2 <- repeat_2024_merged_2 %>% mutate(Time_body_out_sec_raw = ifelse(is.na(Time_body_out_sec_raw), time_limit, Time_body_out_sec_raw))

# Merge data frames
merged_repeat_2024 <- bind_rows(repeat_ind_2024_merged, repeat_2024_merged_2)

merged_repeat_2024 <- merged_repeat_2024[order(merged_repeat_2024$Snake_ID), ]

# Check that all data over 3000s is correctly censored
for(i in 1:nrow(merged_repeat_2024)){
  
  if(merged_repeat_2024$Time_head_out_sec_raw[i] > time_limit){
    
    merged_repeat_2024$Time_head_out_sec_raw[i] <- time_limit
    
    merged_repeat_2024$Time_body_out_sec_raw[i] <- time_limit
    
  } else { }
  
}

```

### Head-out

```{r rep headout 2023-2024}
#| warning: false

# General ICC
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024[, c("Snake_ID", "Time_head_out_sec_raw", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# Divide into CORE and FRONT
merged_repeat_2024_core <- merged_repeat_2024[merged_repeat_2024$Inv_situation == "CORE", ]

merged_repeat_2024_front <- merged_repeat_2024[merged_repeat_2024$Inv_situation == "FRONT", ]

# CORE
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_core[, c("Snake_ID", "Time_head_out_sec_raw", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# FRONT
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_front[, c("Snake_ID", "Time_head_out_sec_raw", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = merged_repeat_2024, aes(x = Repeatability, y = log1p(Time_head_out_sec_raw))) +
  theme_bw() + 
  facet_wrap(~ Inv_situation, labeller = labeller(Inv_situation = 
                                          c("CORE" = "CORE\nICC = 0.368 | p = 0.0623",
                                            "FRONT" = "FRONT\nICC = 0.678 | p < 0.001"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Time to head-out (s))") + 
  ggtitle("Head-out time (s) since start of the experiment (log)") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# CORE
anova_core <- aov(log(Time_head_out_sec_raw) ~ Repeatability, data = merged_repeat_2024_core)

summary(anova_core)

# FRONT
anova_front <- aov(log(Time_head_out_sec_raw) ~ Repeatability, data = merged_repeat_2024_front)

summary(anova_front)
```

### Body-out

```{r rep bodyout 2023-2024}
#| warning: false

# General ICC
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024[, c("Snake_ID", "Time_body_out_sec_raw", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# Divide into CORE and FRONT
merged_repeat_2024_core <- merged_repeat_2024[merged_repeat_2024$Inv_situation == "CORE", ]

merged_repeat_2024_front <- merged_repeat_2024[merged_repeat_2024$Inv_situation == "FRONT", ]

# CORE
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_core[, c("Snake_ID", "Time_body_out_sec_raw", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# FRONT
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_front[, c("Snake_ID", "Time_body_out_sec_raw", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = merged_repeat_2024, aes(x = Repeatability, y = log1p(Time_body_out_sec_raw))) +
  theme_bw() + 
  facet_wrap(~ Inv_situation, labeller = labeller(Inv_situation = 
                                          c("CORE" = "CORE\nICC = 0.342 | p = 0.0777",
                                            "FRONT" = "FRONT\nICC = 0.694 | p < 0.001"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Time to body-out (s))") + 
  ggtitle("Body-out time (s) since start of the experiment (log)") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# CORE
anova_core <- aov(log(Time_body_out_sec_raw) ~ Repeatability, data = merged_repeat_2024_core)

summary(anova_core)

# FRONT
anova_front <- aov(log(Time_body_out_sec_raw) ~ Repeatability, data = merged_repeat_2024_front)

summary(anova_front)
```

### Visual exploration

```{r rep visexp 2023-2024}
#| warning: false

# Delete rows (individuals) that didn't do body-out under 3000 sec (only for visual exploration)
merged_repeat_2024_outside <- subset(merged_repeat_2024, merged_repeat_2024[["Time_body_out_sec_raw"]] != time_limit)

# Detect those individuals that in both their trials, exited the initial refuge
duplicated_outside <- duplicated(merged_repeat_2024_outside[["Snake_ID"]]) | duplicated(merged_repeat_2024_outside[["Snake_ID"]], fromLast = TRUE)

# Select only those individuals
merged_repeat_2024_outside <- merged_repeat_2024_outside[duplicated_outside, ]

# New variable: exploration time
merged_repeat_2024_outside$Visual_exp <- merged_repeat_2024_outside$Time_body_out_sec_raw - merged_repeat_2024_outside$Time_head_out_sec_raw

# General ICC
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_outside[, c("Snake_ID", "Visual_exp", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# Divide into CORE and FRONT
merged_repeat_2024_core <- merged_repeat_2024_outside[merged_repeat_2024_outside$Inv_situation == "CORE", ]

merged_repeat_2024_front <- merged_repeat_2024_outside[merged_repeat_2024_outside$Inv_situation == "FRONT", ]

# CORE
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_core[, c("Snake_ID", "Visual_exp", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# FRONT
# We extract the variable that we want to test repeatability
icc_data <- as.data.frame(merged_repeat_2024_front[, c("Snake_ID", "Visual_exp", "repeat_trial")])

# Put it in wide format
icc_data_wide <- reshape(icc_data, idvar = "Snake_ID", timevar = "repeat_trial", direction = "wide")

# Delete ID column
icc_data_wide <- icc_data_wide[, -1]

# Change column names
colnames(icc_data_wide) <- c("First_trial", "Second_trial")

# log transform varaibles
icc_data_wide <- log1p(icc_data_wide)

# Consistency (ICC)
irr::icc(icc_data_wide, model = "oneway", type = "consistency", unit = "single")

# Plot
ggplot(data = merged_repeat_2024_outside, aes(x = Repeatability, y = log1p(Visual_exp))) +
  theme_bw() + 
  facet_wrap(~ Inv_situation, labeller = labeller(Inv_situation = 
                                          c("CORE" = "CORE\nICC = 0.369 | p = 0.122",
                                            "FRONT" = "FRONT\nICC = 0.291 | p = 0.17"))) + 
  stat_summary(
    geom = "errorbar", 
    fun.data = mean_cl_boot, 
    width = 0.1,
    size = 0.8,  
    col = "black",
    position = position_nudge(x = c(-0.2, 0.2))) +
  geom_point(size = 3) +
  stat_summary(fun.y = mean, position = position_nudge(x = c(-0.2, 0.2)), size = 1) +
  stat_summary(aes(group = 1), fun.y = mean, geom = "line", size = 1.5, linetype = "solid", position = position_nudge(x = c(-0.2, 0.2))) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "pointrange", color = "black", size = 0.2,
               position = position_nudge(x = c(-0.2, 0.2))) +
  geom_line(aes(group = Snake_ID), size = 0.5, linetype = "dashed") + 
  ylab("log(Visual exploration time (s))") + 
  ggtitle("Visual exploration time (s) since start of the experiment (log)") +
  theme(plot.caption = element_text(hjust = 0.5))

# ANOVA differences in mean
# CORE
anova_core <- aov(log(Visual_exp) ~ Repeatability, data = merged_repeat_2024_core)

summary(anova_core)

# FRONT
anova_front <- aov(log(Visual_exp) ~ Repeatability, data = merged_repeat_2024_front)

summary(anova_front)
```

# Survival analysis + Cox regression

Here we performed a survival analysis and a Cox regression to see if there are differences between snakes from the invasion front (*FRONT*) and those from the initially invaded zone (*CORE*) for the time it takes for the snakes to get their heads out of the initial refuge for the first time (*Head-out*) and to get completely out of the initial refuge for the first time (*Body-out*). This analysis doesn't include 2024 snakes.

## No log-transformation

```{r survival analysis}
#| warning: false
# We add columns for saying if data is censored (0) or the event has happened within the time frame (1)
summary_tracking_complete_no_rep <- add_column(summary_tracking_complete_no_rep, Head_out_01 = NA, .after = "Time_head_out_sec_raw")

summary_tracking_complete_no_rep <- add_column(summary_tracking_complete_no_rep, Body_out_01 = NA, .after = "Time_body_out_sec_raw")

# 1 for event (head-out/body-out) and 0 for censored data
for(i in 1:nrow(summary_tracking_complete_no_rep)){
  
  if(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] == 3000){
    
    summary_tracking_complete_no_rep$Head_out_01[i] <- 0
    
  } else {
    
    summary_tracking_complete_no_rep$Head_out_01[i] <- 1
    
  }
  
  if(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] == 3000){
    
    summary_tracking_complete_no_rep$Body_out_01[i] <- 0
    
  } else {
    
    summary_tracking_complete_no_rep$Body_out_01[i] <- 1
    
  }
  
}

# Head-out
fit_head_out <- survfit(Surv(Time_head_out_sec_raw, Head_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)

ggsurvplot(fit_head_out,
           pval = TRUE,
           conf.int = TRUE,
           risk.table = TRUE, 
           risk.table.col = "strata", 
           linetype = c(1, 1),
           ggtheme = theme_bw(),
           palette = c("#E7B800", "#2E9FDF"),
           surv.median.line = "hv",
           legend.labs = c("CORE", "FRONT"))

coxph(Surv(Time_head_out_sec_raw, Head_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)

# Body-out
fit_body_out <- survfit(Surv(Time_body_out_sec_raw, Body_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)

ggsurvplot(fit_body_out,
           pval = TRUE,
           conf.int = TRUE,
           risk.table = TRUE, 
           risk.table.col = "strata", 
           linetype = c(1, 1),
           ggtheme = theme_bw(),
           palette = c("#E7B800", "#2E9FDF"),
           surv.median.line = "hv",
           legend.labs = c("CORE", "FRONT"))

coxph(Surv(Time_body_out_sec_raw, Body_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)
```

## Log-transformation

```{r log survival analysis}
# Head-out
fit_head_out <- survfit(Surv(log(Time_head_out_sec_raw), Head_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)

ggsurvplot(fit_head_out,
           pval = TRUE,
           conf.int = TRUE,
           risk.table = TRUE, 
           risk.table.col = "strata", 
           linetype = c(1, 1),
           ggtheme = theme_bw(),
           palette = c("#E7B800", "#2E9FDF"),
           surv.median.line = "hv",
           legend.labs = c("CORE", "FRONT"))

coxph(Surv(log(Time_head_out_sec_raw), Head_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)

# Body-out
fit_body_out <- survfit(Surv(log(Time_body_out_sec_raw), Body_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)

ggsurvplot(fit_body_out,
           pval = TRUE,
           conf.int = TRUE,
           risk.table = TRUE, 
           risk.table.col = "strata", 
           linetype = c(1, 1),
           ggtheme = theme_bw(),
           palette = c("#E7B800", "#2E9FDF"),
           surv.median.line = "hv",
           legend.labs = c("CORE", "FRONT"))

coxph(Surv(log(Time_body_out_sec_raw), Body_out_01) ~ Inv_situation, data = summary_tracking_complete_no_rep)
```

# Censored data regressions

In this section, we will use a method for estimating linear regressions with censored data. For the moment, I am including gravid females, but they can be removed.

This part of the analysis has been done with all sexes at the same time, to determine the power of the analyses with censored data, and to see if the results of a normal analysis, one with censored data, and one without censored data are very different from each other. Ideally, it should be separated by gender, for example.

## Head-out

### No log-transformation

```{r censored head out}
#| warning: false

# We create a data frame for only the "OLS detect" linear regression
summary_tracking_complete_no_rep_detect <- summary_tracking_complete_no_rep[
  summary_tracking_complete_no_rep$Time_head_out_sec_raw >= 0 & 
  summary_tracking_complete_no_rep$Time_head_out_sec_raw < 3000, 
]

# ~ Invasion year
# Censored regression
fit.cen <- censReg((Time_head_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep, left = -Inf, right = (3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Time_head_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Time_head_out_sec_raw ~ Year_invasion, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Year_invasion, (summary_tracking_complete_no_rep$Time_head_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = 'Head-out vs. Invasion year', xlab = "Invasion year", ylab = "Time (s)")

lineplot()

# ~ Distance to Noah's garden
# Censored regression
fit.cen <- censReg((Time_head_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep, left = -Inf, right = (3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Time_head_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Time_head_out_sec_raw ~ Distance_Noahs, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Distance_Noahs, (summary_tracking_complete_no_rep$Time_head_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "Head-out vs. Distance to Noah's garden", xlab = "Distance to Noah's garden (m)", ylab = "Time (s)")

lineplot()


# ~ Condition
# Censored regression
fit.cen <- censReg((Time_head_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep, left = -Inf, right = (3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Time_head_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Time_head_out_sec_raw ~ Condition, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Condition, (summary_tracking_complete_no_rep$Time_head_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "Head-out vs. Condition", xlab = "Condition", ylab = "Time (s)")

lineplot()

```

### Log-transformation

```{r log censored head out}
#| warning: false
#| # ~ Invasion year
# Censored regression
fit.cen <- censReg(log(Time_head_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep, left = -Inf, right = log(3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm(log(Time_head_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(log(Time_head_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Year_invasion, log(summary_tracking_complete_no_rep$Time_head_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "log(Head-out) vs. Invasion year", xlab = "Invasion year", ylab = "log(Time (s))")

lineplot()

# ~ Distance to Noah's garden
# Censored regression
fit.cen <- censReg(log(Time_head_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep, left = -Inf, right = log(3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm(log(Time_head_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(log(Time_head_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Distance_Noahs, log(summary_tracking_complete_no_rep$Time_head_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "log(Head-out) vs. Distance to Noah's garden", xlab = "Distance to Noah's garden (m)", ylab = "log(Time (s))")

lineplot()


# ~ Condition
# Censored regression
fit.cen <- censReg(log(Time_head_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep, left = -Inf, right = log(3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm(log(Time_head_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(log(Time_head_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Condition, log(summary_tracking_complete_no_rep$Time_head_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "log(Head-out) vs. Condition", xlab = "Condition", ylab = "log(Time (s))")

lineplot()

```

## Body-out

### No log-transformation

```{r censored body out}
#| warning: false
# ~ Invasion year
# Censored regression
fit.cen <- censReg((Time_body_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep, left = -Inf, right = (3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Time_body_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm((Time_body_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Year_invasion, (summary_tracking_complete_no_rep$Time_body_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "Body-out vs. Invasion year", xlab = "Invasion year", ylab = "Time (s)")

lineplot()

# ~ Distance to Noah's garden
# Censored regression
fit.cen <- censReg((Time_body_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep, left = -Inf, right = (3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Time_body_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm((Time_body_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Distance_Noahs, (summary_tracking_complete_no_rep$Time_body_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "Body-out vs. Distance to Noah's garden", xlab = "Distance to Noah's garden (m)", ylab = "Time (s)")

lineplot()


# ~ Condition
# Censored regression
fit.cen <- censReg((Time_body_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep, left = -Inf, right = (3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Time_body_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm((Time_body_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Condition, (summary_tracking_complete_no_rep$Time_body_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "Body-out vs. Condition", xlab = "Condition", ylab = "Time (s)")

lineplot()

```

### Log-transformation

```{r log censored body out}
#| warning: false
#| # ~ Invasion year
# Censored regression
fit.cen <- censReg(log(Time_body_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep, left = -Inf, right = log(3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm(log(Time_body_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(log(Time_body_out_sec_raw) ~ Year_invasion, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Year_invasion, log(summary_tracking_complete_no_rep$Time_body_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "log(Body-out) vs. Invasion year", xlab = "Invasion year", ylab = "log(Time (s))")

lineplot()

# ~ Distance to Noah's garden
# Censored regression
fit.cen <- censReg(log(Time_body_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep, left = -Inf, right = log(3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm(log(Time_body_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(log(Time_body_out_sec_raw) ~ Distance_Noahs, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Distance_Noahs, log(summary_tracking_complete_no_rep$Time_body_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "log(Body-out) vs. Distance to Noah's garden", xlab = "Distance to Noah's garden (m)", ylab = "log(Time (s))")

lineplot()


# ~ Condition
# Censored regression
fit.cen <- censReg(log(Time_body_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep, left = -Inf, right = log(3000))

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm(log(Time_body_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(log(Time_body_out_sec_raw) ~ Condition, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Condition, log(summary_tracking_complete_no_rep$Time_body_out_sec_raw), pch = 19, cex = 0.7, col = 'grey60', main = "log(Body-out) vs. Condition", xlab = "Condition", ylab = "log(Time (s))")

lineplot()
```

## Visual exploration

### No log-transformation

```{r censored visual}
#| warning: false

# Create filtered data frame for this part of the analysis
summary_tracking_complete_no_rep_detect <- summary_tracking_complete_no_rep[summary_tracking_complete_no_rep$Visual_exp != 0, ]

# ~ Invasion year
# Censored regression
fit.cen <- censReg((Visual_exp) ~ Year_invasion, data = summary_tracking_complete_no_rep, left = 0, right = Inf)

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Visual_exp) ~ Year_invasion, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Visual_exp ~ Year_invasion, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS *",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Year_invasion, (summary_tracking_complete_no_rep$Visual_exp), pch = 19, cex = 0.7, col = 'grey60', main = "Visual exploration vs. Invasion year", xlab = "Invasion year", ylab = "Time (s)")

lineplot()

# ~ Distance to Noah's garden
# Censored regression
fit.cen <- censReg((Visual_exp) ~ Distance_Noahs, data = summary_tracking_complete_no_rep, left = 0, right = Inf)

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Visual_exp) ~ Distance_Noahs, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Visual_exp ~ Distance_Noahs, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored *",
      "OLS *",
      "OLS Detects *"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Distance_Noahs, (summary_tracking_complete_no_rep$Visual_exp), pch = 19, cex = 0.7, col = 'grey60', main = "Visual exploration vs. Distance to Noah's garden", xlab = "Distance to Noah's garden (m)", ylab = "Time (s)")

lineplot()


# ~ Condition
# Censored regression
fit.cen <- censReg((Visual_exp) ~ Condition, data = summary_tracking_complete_no_rep, left = 0, right = Inf)

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Visual_exp) ~ Condition, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Visual_exp ~ Condition, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Condition, (summary_tracking_complete_no_rep$Visual_exp), pch = 19, cex = 0.7, col = 'grey60', main = "Visual exploration vs. Condition", xlab = "Condition", ylab = "Time (s)")

lineplot()

```

### Log-transformation

```{r log censored visual}
#| warning: false

# Create log variable
summary_tracking_complete_no_rep$Visual_exp_log <- log1p(summary_tracking_complete_no_rep$Visual_exp)

# Create the same variable for the only detection data frame
summary_tracking_complete_no_rep_detect$Visual_exp_log <- log1p(summary_tracking_complete_no_rep_detect$Visual_exp)

# ~ Invasion year
# Censored regression
fit.cen <- censReg((Visual_exp_log) ~ Year_invasion, data = summary_tracking_complete_no_rep, left = 0, right = Inf)

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Visual_exp_log) ~ Year_invasion, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Visual_exp_log ~ Year_invasion, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Year_invasion, (summary_tracking_complete_no_rep$Visual_exp_log), pch = 19, cex = 0.7, col = 'grey60', main = "log(Visual exploration) vs. Invasion year", xlab = "Invasion year", ylab = "log(Time (s))")

lineplot()

# ~ Distance to Noah's garden
# Censored regression
fit.cen <- censReg((Visual_exp_log) ~ Distance_Noahs, data = summary_tracking_complete_no_rep, left = 0, right = Inf)

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Visual_exp_log) ~ Distance_Noahs, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Visual_exp_log ~ Distance_Noahs, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects *"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Distance_Noahs, (summary_tracking_complete_no_rep$Visual_exp_log), pch = 19, cex = 0.7, col = 'grey60', main = "log(Visual exploration) vs. Distance to Noah's garden", xlab = "Distance to Noah's garden (m)", ylab = "log(Time (s))")

lineplot()


# ~ Condition
# Censored regression
fit.cen <- censReg((Visual_exp_log) ~ Condition, data = summary_tracking_complete_no_rep, left = 0, right = Inf)

summary(fit.cen)

# Fit a line using OLS regression (linear regression)
fit.OLS <- lm((Visual_exp_log) ~ Condition, data = summary_tracking_complete_no_rep)

summary(fit.OLS)

# fit regression only of non-censored data
fit.detect <- lm(Visual_exp_log ~ Condition, data = summary_tracking_complete_no_rep_detect)

summary(fit.detect)

# Plot the results
lineplot <- function() {
  
  # Add lines
  abline(coef(fit.cen)[1:2], col = "red", lwd = 2)
  abline(coef(fit.OLS), col = "blue", lty = 2, lwd = 2)
  abline(coef(fit.detect), col = "green", lty = 3, lwd = 2)
  
  # Add legend
  legend(
    "topleft",
    legend = c(
      "Censored",
      "OLS",
      "OLS Detects"
    ),
    lwd = 3 * par("cex"),
    col = c("red", "blue", "green"),
    lty = c(1, 2, 3),
    text.font = 1, bty = "n",
    pt.cex = 1, cex = 0.8, y.intersp = 1.3
  )
}

plot(summary_tracking_complete_no_rep$Condition, (summary_tracking_complete_no_rep$Visual_exp_log), pch = 19, cex = 0.7, col = 'grey60', main = "log(Visual exploration) vs. Condition", xlab = "Condition", ylab = "log(Time (s))")

lineplot()
```

# PERMANOVA + betadisper

This section analyses whether there are differences in the behaviour and use of the snake area. The analyses contain gravid females, but they can be removed if necessary.

## Correlation analysis

First we are going to carry out a correlation test between variables to see if we can omit some of them when carrying out the analyses, as they could be highly correlated with each other.

```{r correlation}
#| warning: false

# Select environmental variables
variables_lm <- summary_tracking_complete_no_rep[, c("Distance_Noahs", "Year_invasion", "SVL", "Condition", "Temperature", "R_H")]

# Transform SVL to log
variables_lm$SVL <- log(variables_lm$SVL)

# Correlation results
cor(variables_lm)

# Correlation plot
corrplot(cor(variables_lm), method="color")

```

Distance to Noah's garden and year of invasion are highly correlated with each other, as are SVL and condition, or relative humidity and temperature. For the analyses to come, we will use only one of each pair of these correlations to perform the analyses (distance to Noah's garden, SVL, and temperature), along with the sex variable.

## Behaviour analysis

In this section we are going to analyse whether for a given study time (15 min, for example), we observe differences in the type of behaviours expressed by the snakes, depending on different variables (PERMANOVA), and whether the variability of these behaviours is greater or lesser in certain groups (betadisper).

### Stacked behaviour plot 

```{r stacked behaviour}
#| warning: false

# Delete duplications from the array
# No repeatability filter

# Duplicate data frame
Snake_array_dupl <- Snake_array

# No repeatability filter
snake_id <- names(Snake_array[1,1,])

dupl_positions <- array(data = NA)

# Identify duplicated positions
a <- 1

for (i in 1:length(snake_id)){
  
  if(duplicated(snake_id)[i] == FALSE){
    
    
  } else {
    
    dupl_positions[a] <- i
    
    a <- a + 1
    
  }
  
}

# Final data frame without duplicates (repeatability)
Snake_array_dupl <- Snake_array_dupl[, , -dupl_positions]

# How much time do we want to analyse
study_time <- 900

# Create data frame
behaviour_df <- as.data.frame(matrix(data = NA, ncol = 15, nrow = length(names(Snake_array_dupl[1,1,]))))

# Column names
colnames(behaviour_df) <- c("Snake_ID", "headout_bf_bodyout", "headout_after_bodyout", "exploring_arena", "exploring_wall", "immobile_exposed", "hiding", "hiding_corner", "year", "distance", "Sex", "SVL", "treatment", "n_eggs", "temperature")

# For all individuals
for (i in 1:length(names(Snake_array_dupl[1,1,]))){

  # If the individual put its head out of the initial refuge 15 min (study time) before the time ended:
  if(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time <= 3000){
    
    # ID individual
    behaviour_df$Snake_ID[i] <- summary_tracking_complete_no_rep$Snake_ID[i]
    
    # How much time did the snake use doing head-out behaviour(visual exploration) before exiting the initial refuge for the first time
    behaviour_df$headout_bf_bodyout[i] <- sum(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + 1):min((summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time), summary_tracking_complete_no_rep$Time_body_out_sec_raw[i]), 22, i] == "head-out")
    
    # How much time did the snake do head-out in the rest of refuges after exiting for the first time the initial refuge
    behaviour_df$headout_after_bodyout[i] <- sum(Snake_array_dupl[min((summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1),(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time + 1)):(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time + 1), 22, i] == "head-out")
    
    # How much time does the snake spend moving around the arena
    behaviour_df$exploring_arena[i] <- sum(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time), 22, i] == "moving")
    
    # How much time did the snake spend exploring the walls (vertically)
    behaviour_df$exploring_wall[i] <- sum(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time), 22, i] == "exploring wall")
    
    # How much time did the snake spend inmobile in the midle of the arena
    behaviour_df$immobile_exposed[i] <- sum(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time), 22, i] == "immobile exposed")
    
    # How much time did the snake hide during trial
    behaviour_df$hiding[i] <- sum(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time), 22, i] == "hidden")
    
    behaviour_df$hiding_corner[i] <- sum(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_head_out_sec_raw[i] + study_time), 22, i] == "hiding / head-out")
    
    # Other variables
    behaviour_df$year[i] <- summary_tracking_complete_no_rep$Year_invasion[i]
    
    behaviour_df$distance[i] <- summary_tracking_complete_no_rep$Distance_Noahs[i]
    
    behaviour_df$Sex[i] <- summary_tracking_complete_no_rep$Sex[i]
    
    behaviour_df$SVL[i] <- summary_tracking_complete_no_rep$SVL[i]
    
    behaviour_df$treatment[i] <- summary_tracking_complete_no_rep$Inv_situation[i]
    
    behaviour_df$n_eggs[i] <- summary_tracking_complete_no_rep$N_eggs[i]
    
    behaviour_df$temperature[i] <- summary_tracking_complete_no_rep$Temperature[i]
    
    
  } else {
    
    
  }
  
}


behaviour_df <- behaviour_df %>% drop_na(Snake_ID)

# Behaviour
# Select columns of interest for stacked behaviour plot
stacked_behaviour <- behaviour_df[, c("Snake_ID", "headout_bf_bodyout", "headout_after_bodyout", "exploring_arena", "exploring_wall", "immobile_exposed", "hiding", "hiding_corner")]

# Snake_ID ordered by year of invasion
snakes_order_by_distance <- behaviour_df[, c("Snake_ID", "distance")]

snakes_order_by_distance <- snakes_order_by_distance[order(snakes_order_by_distance$distance, decreasing = FALSE), ]

# Long format
stacked_behaviour_long <- pivot_longer(stacked_behaviour, cols = c("headout_bf_bodyout", "headout_after_bodyout", "exploring_arena", "exploring_wall", "immobile_exposed", "hiding", "hiding_corner"), names_to = "Behaviour", values_to = "Percentage")

# Percentage of each behaviour
stacked_behaviour_long <- stacked_behaviour_long %>%
  group_by(Snake_ID) %>%
  mutate(Percentage_total = sum(Percentage)) %>%
  ungroup() %>%
  mutate(Percentage = Percentage / Percentage_total * 100)

# Order Snake ID by year of invasion
stacked_behaviour_long$Snake_ID <- factor(stacked_behaviour_long$Snake_ID, levels = snakes_order_by_distance$Snake_ID)

# Order behaviour variables
stacked_behaviour_long$Behaviour <- factor(stacked_behaviour_long$Behaviour, levels = c("immobile_exposed", "exploring_wall", "exploring_arena",  "headout_after_bodyout", "headout_bf_bodyout", "hiding_corner", "hiding"))

# Colors for behaviours
beh_colors <- c(viridis(7)[7], viridis(7)[6], viridis(7)[5], viridis(7)[4], viridis(7)[3], viridis(7)[2], viridis(7)[1])

# Plot
ggplot(stacked_behaviour_long, aes(x = Snake_ID, y = Percentage, fill = Behaviour)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Snake ID", y = paste("% behaviours during", study_time/60, "min test"), fill = "Behaviours") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10), 
        legend.text = element_text(size = 10), 
        axis.title.x = element_text(size = 15),  
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 10), 
        legend.title = element_text(size = 15)) +
  scale_fill_manual(values = beh_colors)


```

### PERMANOVA

```{r PERMANOVA behaviour}
#| warning: false

# Alternative data frame with invasion situation
stacked_behaviour_treatment <- behaviour_df

# Delete Snake ID column
stacked_behaviour_complete <- stacked_behaviour_treatment

stacked_behaviour_treatment <- subset(stacked_behaviour_treatment, select = -c(Snake_ID))

# Subset pca columns (numeric)
stacked_behaviour_treatment_var <- stacked_behaviour_treatment[1:7]

# PCA
pca_beh <- prcomp(stacked_behaviour_treatment_var, scale. = T)

# Weights PCA axis
pca_beh$rotation

# Summary
summary(pca_beh)

# Plot front vs. core
fviz_pca_biplot(pca_beh, geom.ind = "point", 
             habillage = stacked_behaviour_treatment$treatment, 
             axes = c(1, 2), 
             pointsize = 3,
             addEllipses = TRUE, ellipse.level = 0.9) 

# Plot sex
fviz_pca_biplot(pca_beh, geom.ind = "point", 
             habillage = stacked_behaviour_treatment$Sex, 
             axes = c(1, 2), 
             pointsize = 3,
             addEllipses = TRUE, ellipse.level = 0.9) 

# PERMANOVA variables
stacked_behaviour_treatment_env <- stacked_behaviour_treatment[, c("treatment", "distance", "SVL", "Sex", "temperature")]

# PERMANOVA without interaction
beh.permanova <- adonis2(stacked_behaviour_treatment_var ~ distance + SVL + Sex + temperature, data = stacked_behaviour_treatment_env, permutations = 999, method="euclidean")

beh.permanova

# PERMANOVA with interaction
beh.permanova <- adonis2(stacked_behaviour_treatment_var ~ distance * SVL * Sex * temperature, data = stacked_behaviour_treatment_env, permutations = 999, method="euclidean")

beh.permanova
```

### betadisper

```{r betadisper behaviour}
#| warning: false

# Front vs. core
# Calculate distance matrix
beh.dist <- vegdist(decostand(stacked_behaviour_treatment_var, "norm"), method="euclidean")

beh.dispersion <- betadisper(beh.dist, group = stacked_behaviour_treatment_env$treatment)

permutest(beh.dispersion)

par(mfrow = c(1, 2))

plot(beh.dispersion, hull = FALSE, ellipse = TRUE, main = "PCoA front vs. core dispersion") 

boxplot(beh.dispersion, main = "Boxplot distance to centroids", xlab = "Groups")

# Sex
# Calculate distance matrix
beh.dist <- vegdist(decostand(stacked_behaviour_treatment_var, "norm"), method="euclidean")

beh.dispersion <- betadisper(beh.dist, group = stacked_behaviour_treatment_env$Sex)

permutest(beh.dispersion)

par(mfrow = c(1, 2))

plot(beh.dispersion, hull = FALSE, ellipse = TRUE, main = "PCoA M vs. F dispersion") 

boxplot(beh.dispersion, main = "Boxplot distance to centroids", xlab = "Groups")

```

## Area use analysis

In contrast to behaviour, here we want to know how much time did each snake spend doing head-out before and after body-out. We are going to analyse the use of the area only after body-out, because if we analysed area use after head-out, the initial refuge area would have bigger numbers. We want to know what areas did the snake visit after exiting the refuge, not taking into account the initial visual exploration that snakes performed during head-out before body-out.

### Stacked area use plot

```{r stacked area}
#| warning: false

# Duplicate original data frame
Snake_array_dupl <- Snake_array

# No repeatability filter
snake_id <- names(Snake_array[1,1,])

dupl_positions <- array(data = NA)

a <- 1

for (i in 1:length(snake_id)){
  
  if(duplicated(snake_id)[i] == FALSE){
    
    
  } else {
    
    dupl_positions[a] <- i
    
    a <- a + 1
    
  }
  
}

# Final data frame without duplicates (repeatability)
Snake_array_dupl <- Snake_array_dupl[, , -dupl_positions]

# How much time do we want to analyse
study_time <- 900

# Create data frame
area_df <- as.data.frame(matrix(data = NA, ncol = 16, nrow = length(names(Snake_array_dupl[1,1,]))))

# Column names
colnames(area_df) <- c("Snake_ID", "time_central_area", "time_intermediate_area", "time_refuge_area_1", "time_refuge_area_2", "time_initial_refuge", "time_distant_wall", "time_lateral_1", "time_lateral_2", "year", "distance", "Sex", "SVL", "treatment", "n_eggs", "temperature")

# For all snakes:
for (i in 1:length(names(Snake_array_dupl[1,1,]))){
  
  # If the snake exited the initial refuge and explored for more than the specified syudy time before time ran out
  if(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time <= 3000){
    
    # Snake ID
    area_df$Snake_ID[i] <- summary_tracking_complete_no_rep$Snake_ID[i]
    
    # Time spent in each of the areas
    area_df$time_central_area[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Central area"))
    
    area_df$time_intermediate_area[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Intermediate area"))
    
    area_df$time_refuge_area_1[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Refuge area 1"))
    
    area_df$time_refuge_area_2[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Refuge area 2"))
    
    area_df$time_initial_refuge[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Initial refuge"))
    
    area_df$time_distant_wall[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Distant wall"))
    
    area_df$time_lateral_1[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Lateral 1"))
    
    area_df$time_lateral_2[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 13, i] == "Lateral 2"))
    
    # Other variables
    area_df$year[i] <- summary_tracking_complete_no_rep$Year_invasion[i]
    
    area_df$distance[i] <- summary_tracking_complete_no_rep$Distance_Noahs[i]
    
    area_df$Sex[i] <- summary_tracking_complete_no_rep$Sex[i]
    
    area_df$SVL[i] <- summary_tracking_complete_no_rep$SVL[i]
    
    area_df$treatment[i] <- summary_tracking_complete_no_rep$Inv_situation[i]
    
    area_df$n_eggs[i] <- summary_tracking_complete_no_rep$N_eggs[i]
    
    area_df$temperature[i] <- summary_tracking_complete_no_rep$Temperature[i]
    
    
  } else {
    
    
  }
  
}


area_df <- area_df %>% drop_na(Snake_ID)


#Select columns
stacked_area <- area_df[, c("Snake_ID", "time_central_area", "time_intermediate_area", "time_refuge_area_1", "time_refuge_area_2", "time_initial_refuge", "time_distant_wall", "time_lateral_1", "time_lateral_2")]

# Snake_ID ordered by distance to Noah's garden
snakes_order_by_distance <- area_df[, c("Snake_ID", "distance")]

snakes_order_by_distance <- snakes_order_by_distance[order(snakes_order_by_distance$distance, decreasing = FALSE), ]

# New variables (merge laterals)
stacked_area$refuge_area <- stacked_area$time_refuge_area_1 + stacked_area$time_refuge_area_2

stacked_area$over_lateral <- stacked_area$time_lateral_1 + stacked_area$time_lateral_2

# We delete these intermediate variables used to calculate previous variables
stacked_area <- subset(stacked_area, select = -c(time_refuge_area_1, time_refuge_area_2, time_lateral_1, time_lateral_2))

# Alternative data frame with invasion situation
stacked_area_treatment <- stacked_area

stacked_area_treatment$treatment <- area_df$treatment

stacked_area_treatment$year <- area_df$year

stacked_area_treatment$SVL <- area_df$SVL

stacked_area_treatment$distance <- area_df$distance

stacked_area_treatment$sex <- area_df$Sex

stacked_area_treatment$temperature <- area_df$temperature

# Long format
stacked_area_long <- pivot_longer(stacked_area, cols = c("time_central_area", "time_intermediate_area", "time_initial_refuge", "time_distant_wall", "refuge_area", "over_lateral"), names_to = "Area", values_to = "Percentage")

# Percentage of each behaviour
stacked_area_long <- stacked_area_long %>%
  group_by(Snake_ID) %>%
  mutate(Percentage_total = sum(Percentage)) %>%
  ungroup() %>%
  mutate(Percentage = Percentage / Percentage_total * 100)

# Order Snake ID by distance to Noah's garden
stacked_area_long$Snake_ID <- factor(stacked_area_long$Snake_ID, levels = snakes_order_by_distance$Snake_ID)

# Order behaviour variables
stacked_area_long$Area <- factor(stacked_area_long$Area, levels = c("time_central_area", "time_intermediate_area", "time_initial_refuge", "time_distant_wall", "refuge_area", "over_lateral"))


# Colors for behaviours
area_colors <- c(viridis(6)[6], viridis(6)[5], viridis(6)[4], viridis(6)[3], viridis(6)[2], viridis(6)[1])

# Plot
ggplot(stacked_area_long, aes(x = Snake_ID, y = Percentage, fill = Area)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Snake ID", y = paste("% area use during", study_time/60, "min test"), fill = "Area") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10), 
        legend.text = element_text(size = 10), 
        axis.title.x = element_text(size = 15),  
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 10), 
        legend.title = element_text(size = 15)) +
  scale_fill_manual(values = area_colors)

```

### PERMANOVA

```{r PERMANOVA area}
#| warning: false

# Delete Snake ID column
stacked_area_complete <- stacked_area_treatment

stacked_area_treatment <- subset(stacked_area_treatment, select = -c(Snake_ID))

# Subset pca columns (numeric)
stacked_area_treatment_var <- stacked_area_treatment[, c("time_central_area", "time_intermediate_area", "time_initial_refuge", "time_distant_wall", "refuge_area", "over_lateral")]

# PCA
pca_area <- prcomp(stacked_area_treatment_var, scale. = T)

# Weights PCA axis
pca_area$rotation

# Summary
summary(pca_area)

# Plot front vs. core
fviz_pca_biplot(pca_area, geom.ind = "point", 
                habillage = stacked_area_treatment$treatment, 
                axes = c(1, 2), 
                pointsize = 3,
                addEllipses = TRUE, ellipse.level = 0.9) 

# Plot sex
fviz_pca_biplot(pca_area, geom.ind = "point", 
                habillage = stacked_area_treatment$sex, 
                axes = c(1, 2), 
                pointsize = 3,
                addEllipses = TRUE, ellipse.level = 0.9)

# PERMANOVA variables
stacked_area_treatment_env <- stacked_area_treatment[, c("treatment", "distance", "SVL", "sex", "temperature")]

# PERMANOVA without interaction
beh.area <- adonis2(stacked_area_treatment_var ~ distance + SVL + sex + temperature, data = stacked_area_treatment_env, permutations = 999, method="euclidean")

beh.area

# PERMANOVA with interaction
beh.area <- adonis2(stacked_area_treatment_var ~ distance * SVL * sex * temperature, data = stacked_area_treatment_env, permutations = 999, method="euclidean")

beh.area
```

### betadisper

```{r betadisper area}
#| warning: false

# Front vs. core
# Calculate distance matrix
area.dist <- vegdist(decostand(stacked_area_treatment_var, "norm"), method="euclidean")

area.dispersion <- betadisper(area.dist, group = stacked_area_treatment_env$treatment)

permutest(area.dispersion)

par(mfrow = c(1, 2))

plot(area.dispersion, hull = FALSE, ellipse = TRUE, main = "PCoA front vs. core dispersion") 

boxplot(area.dispersion, main = "Boxplot distance to centroids", xlab = "Groups")

# Sex
# Calculate distance matrix
area.dist <- vegdist(decostand(stacked_area_treatment_var, "norm"), method="euclidean")

area.dispersion <- betadisper(area.dist, group = stacked_area_treatment_env$sex)

permutest(area.dispersion)

par(mfrow = c(1, 2))

plot(area.dispersion, hull = FALSE, ellipse = TRUE, main = "PCoA M vs. F dispersion") 

boxplot(area.dispersion, main = "Boxplot distance to centroids", xlab = "Groups")

```

# Cumulative exploration tests

In this section we will analyse whether there are differences between groups (FRONT vs. CORE and M vs. F) in terms of the behaviour expressed during a certain exploration time (i. e. 15 min). The variables we will look at are the average speed of the individuals, the total distance travelled (in total, including vertical, and only while exploring), the total time exploring, the total time hidden, the number of visits to refuges, the average time the snake has been hidden in these refuges, and the average time the snake is exposed between visits to refuges.

These results include pregnant females. It could be done without pregnant females as well. These results are also exclusive to a specific time range (15 min in this case), but the total time to be analysed could be modified.

## Data filtering and shared functions

```{r cumulative data filtering + functions}
#| warning: false

# Duplicate data frame
Snake_array_dupl <- Snake_array

# No repeatability filter
snake_id <- names(Snake_array[1,1,])

dupl_positions <- array(data = NA)

a <- 1

for (i in 1:length(snake_id)){
  
  if(duplicated(snake_id)[i] == FALSE){
    
    
  } else {
    
    dupl_positions[a] <- i
    
    a <- a + 1
    
  }
  
}

# Final data frame without duplicates (repeatability)
Snake_array_dupl <- Snake_array_dupl[, , -dupl_positions]

# How much time do we want to analyse
study_time <- 900

# Create data frame
movement_df <- as.data.frame(matrix(data = NA, ncol = 18, nrow = length(names(Snake_array_dupl[1,1,]))))

# Column names
colnames(movement_df) <- c("Snake_ID", "avg_speed", "total_dist_trav", "avg_speed_moving", "total_dist_trav_moving", "time_moving", "time_hidden", "n_visits_refuge", "avg_time_hidden", "avg_time_bw_refuges", "year", "distance", "Sex", "SVL", "Condition", "treatment", "n_eggs", "temperature")

for (i in 1:length(names(Snake_array_dupl[1,1,]))){

  # body-out better than head-out
  if(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time <= 3000){
    
    # snake ID
    movement_df$Snake_ID[i] <- summary_tracking_complete_no_rep$Snake_ID[i]
    
    # avg speed during the tested time, independent of behaviour
    movement_df$avg_speed[i] <- mean(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 21, i]))
    
    # total distance travelled, independent of behaviour
    movement_df$total_dist_trav[i] <- sum(as.numeric(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), 21, i]))
    
    
    # subsample: extract only rows when snake is moving
    
    movement_df_moving <- as.data.frame(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), , i])

    movement_df_moving <- subset(movement_df_moving, behaviour == "moving", select = t:behaviour)
    
    movement_df$avg_speed_moving[i] <- mean(as.numeric(movement_df_moving[, 21]))
    
    movement_df$total_dist_trav_moving[i] <- sum(as.numeric(movement_df_moving[, 21]))
    
    movement_df$time_moving[i] <- sum(movement_df_moving[, 22] == "moving")
    
    # subsample: extract only rows when snake is hidden
    
    movement_df_hidden <- as.data.frame(Snake_array_dupl[(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + 1):(summary_tracking_complete_no_rep$Time_body_out_sec_raw[i] + study_time), , i])
    
    movement_df_hidden <- subset(movement_df_hidden, behaviour == "hidden", select = t:behaviour)
    
    movement_df$time_hidden[i] <- sum(movement_df_hidden[, 22] == "hidden")
    
    # differences between numbers
    diff_hidden <- diff(as.numeric(movement_df_hidden$t))
    
    jumps <- diff_hidden[diff_hidden > 1]
    
    n_jumps <- length(jumps)
    
    jump_index <- which(diff_hidden > 1)
    
    chunk <- split(as.numeric(movement_df_hidden$t), cumsum(c(1, diff_hidden > 1)))
  
    chunk_length <- sapply(chunk, length)

    # jump demonstrates a change from hide 1 to hide 2
    
    # We are going to select only those visits that last for longer than 5 seconds, as visits that last for less than 5 seconds, it's just exploration. The snake doesn't hide completely, it only enters its head and exits in a moment
    valid_visits <- chunk_length > 5
    
    n_jumps_valid <- sum(valid_visits) - 1
    
    movement_df$n_visits_refuge[i] <- n_jumps + 1
    
    movement_df$avg_time_bw_refuges[i] <- mean(jumps[valid_visits[-length(valid_visits)]])
    
    movement_df$avg_time_hidden[i] <- mean(chunk_length[valid_visits])
    
    movement_df$year[i] <- summary_tracking_complete_no_rep$Year_invasion[i]
    
    movement_df$distance[i] <- summary_tracking_complete_no_rep$Distance_Noahs[i]
    
    movement_df$Sex[i] <- summary_tracking_complete_no_rep$Sex[i]
    
    movement_df$SVL[i] <- summary_tracking_complete_no_rep$SVL[i]
    
    movement_df$Condition[i] <- summary_tracking_complete_no_rep$Condition[i]
    
    movement_df$treatment[i] <- summary_tracking_complete_no_rep$Inv_situation[i]
    
    movement_df$n_eggs[i] <- summary_tracking_complete_no_rep$N_eggs[i]
    
    movement_df$temperature[i] <- summary_tracking_complete_no_rep$Temperature[i]
    
    
  } else {
    
    
  }
  
}


movement_df <- movement_df %>% drop_na(Snake_ID)

# We delete juveniles, as for the paired comparisons, you need exactly 2 factors
movement_df <- movement_df[which(movement_df$Sex != "Juv"),]

# t-tests
# Function to compare different exploratory variables depending on front vs. core
compare_core_front <- function(variable) {
  t.test(movement_df[[variable]] ~ movement_df$treatment)
}

# Function to compare different exploratory variables depending on male vs. female
compare_sex <- function(variable) {
  t.test(movement_df[[variable]] ~ movement_df$Sex)
}

# To check normality and homoscedasticity
verify_norm <- function(variable, group) {
  shapiro.test(movement_df[[variable]])  
  leveneTest(movement_df[[variable]] ~ movement_df[[group]])  
}

# Plots
# Boxplot 'core' vs 'front'
boxplot_core_front <- function(variable) {
  ggplot(movement_df, aes(x = treatment, y = .data[[variable]], fill = treatment)) +
    geom_boxplot() +
    labs(title = "Boxplots",
         x = "Group", y = variable) +
    theme_minimal()
}

# Boxplot M vs. F
boxplot_sex <- function(variable) {
  ggplot(movement_df, aes(x = Sex, y = .data[[variable]], fill = Sex)) +
    geom_boxplot() +
    labs(title = "",
         x = "Sex", y = variable) +
    theme_minimal()
}

# Scatterplot 'distance'
scatterplot_distance <- function(variable) {
  ggplot(movement_df, aes(x = distance, y = .data[[variable]])) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(title = "Scatterplots",
         x = "Distance to Noah's garden", y = variable) +
    theme_minimal()
}

# Scatterplot 'year'
scatterplot_year <- function(variable) {
  ggplot(movement_df, aes(x = year, y = .data[[variable]])) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(title = "",
         x = "Year", y = variable) +
    theme_minimal()
}

# Scatterplot 'condition'
scatterplot_condition <- function(variable) {
  ggplot(movement_df, aes(x = Condition, y = .data[[variable]])) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(title = "",
         x = "Condition", y = variable) +
    theme_minimal()
}

combine_plots <- function(variable) {
  # Crear los cinco gráficos utilizando las funciones existentes
  plot1 <- boxplot_core_front(variable)
  plot2 <- boxplot_sex(variable)
  plot3 <- scatterplot_distance(variable)
  plot4 <- scatterplot_year(variable)
  plot5 <- scatterplot_condition(variable)
  
  # Combinar los gráficos en una única figura
  combined_plot <- (plot1 | plot2) / (plot3 | plot4 | plot5)
  
  # Mostrar el gráfico combinado
  combined_plot
}
```

## Paired t-tests + boxplots + linear regressions

### Average speed while exploring

```{r avg speed}
#| warning: false

# t-test + normality & homoscedasticity check
# Select variables to compare
variables <- c("avg_speed_moving")

# Front vs. core
results_core_front <- lapply(variables, compare_core_front)
names(results_core_front) <- variables

# Males vs. females
results_sex <- lapply(variables, compare_sex)
names(results_sex) <- variables

# Verify normality and homoscedasticity
verify_core_front <- lapply(variables, verify_norm, group = "treatment")
verify_sex <- lapply(variables, verify_norm, group = "Sex")

# Print results
results_core_front
results_sex
verify_core_front
verify_sex

# lm
model_distance <- lm(avg_speed_moving ~ distance + Condition + Sex + temperature, data = movement_df)
summary(model_distance)

# Plots
combine_plots(variables)

```

### Total distance traveled while exploring

```{r total distance}
#| warning: false

# t-test + normality & homoscedasticity check
# Select variables to compare
variables <- c("total_dist_trav_moving")

# Front vs. core
results_core_front <- lapply(variables, compare_core_front)
names(results_core_front) <- variables

# Males vs. females
results_sex <- lapply(variables, compare_sex)
names(results_sex) <- variables

# Verify normality and homoscedasticity
verify_core_front <- lapply(variables, verify_norm, group = "treatment")
verify_sex <- lapply(variables, verify_norm, group = "Sex")

# Print results
results_core_front
results_sex
verify_core_front
verify_sex

# lm
model_distance <- lm(total_dist_trav_moving ~ distance + Condition + Sex + temperature, data = movement_df)
summary(model_distance)

# Plots
combine_plots(variables)
```

### Number of refuge visits

```{r refuge visits}
#| warning: false

# t-test + normality & homoscedasticity check
# Select variables to compare
variables <- c("n_visits_refuge")

# Front vs. core
results_core_front <- lapply(variables, compare_core_front)
names(results_core_front) <- variables

# Males vs. females
results_sex <- lapply(variables, compare_sex)
names(results_sex) <- variables

# Verify normality and homoscedasticity
verify_core_front <- lapply(variables, verify_norm, group = "treatment")
verify_sex <- lapply(variables, verify_norm, group = "Sex")

# Print results
results_core_front
results_sex
verify_core_front
verify_sex

# lm
model_distance <- lm(n_visits_refuge ~ distance + Condition + Sex + temperature, data = movement_df)
summary(model_distance)

# Plots
combine_plots(variables)
```

### Average time hiding 

```{r avg time hidden}
#| warning: false

# t-test + normality & homoscedasticity check
# Select variables to compare
variables <- c("avg_time_hidden")

# Front vs. core
results_core_front <- lapply(variables, compare_core_front)
names(results_core_front) <- variables

# Males vs. females
results_sex <- lapply(variables, compare_sex)
names(results_sex) <- variables

# Verify normality and homoscedasticity
verify_core_front <- lapply(variables, verify_norm, group = "treatment")
verify_sex <- lapply(variables, verify_norm, group = "Sex")

# Print results
results_core_front
results_sex
verify_core_front
verify_sex

# lm
model_distance <- lm(avg_time_hidden ~ distance + Condition + Sex + temperature, data = movement_df)
summary(model_distance)

# Plots
combine_plots(variables)
```

### Average time between refuge visits

```{r avg time bw refuges}
#| warning: false

# t-test + normality & homoscedasticity check
# Select variables to compare
variables <- c("avg_time_bw_refuges")

# Front vs. core
results_core_front <- lapply(variables, compare_core_front)
names(results_core_front) <- variables

# Males vs. females
results_sex <- lapply(variables, compare_sex)
names(results_sex) <- variables

# Verify normality and homoscedasticity
verify_core_front <- lapply(variables, verify_norm, group = "treatment")
verify_sex <- lapply(variables, verify_norm, group = "Sex")

# Print results
results_core_front
results_sex
verify_core_front
verify_sex

# lm
model_distance <- lm(avg_time_bw_refuges ~ distance + Condition + Sex + temperature, data = movement_df)
summary(model_distance)

# Plots
combine_plots(variables)
```
